---
title: "Exp160_AYEH_241108_FINAL"
author: "Albert Yeh"
date: "2024-11-08"
output: html_document
---
## Used module R/4.4.0 (as of 11/08/24) #Try to use the same version as all dependencies (e.g. package installations) rely on it! This was used from Exp649 (Motoko) was well.


## Output from this experiment:  2 carts; Cart1 and Cart 2.  Example output for Cart 1:
# Cart1_RSEC_ReadsPerCell_Unfiltered.csv.gz
# Cart1_RSEC_ReadsPerCell.csv
# Cart1_RSEC_MolsPerCell_Unfiltered.csv.gz
# Cart1_RSEC_MolsPerCell.csv
# Cart1_DBEC_ReadsPerCell_Unfiltered.csv.gz
# Cart1_DBEC_ReadsPerCell.csv
# Cart1_DBEC_MolsPerCell_Unfiltered.csv.gz
# Cart1_DBEC_MolsPerCell.csv
## BD Handkbook page 17: To remove the effect of UMI errors on molecule counting, BD Biosciences has developed a set of UMI adjustment algorithms.  UMI errors that are single base substitution errors are identified and adjusted to the parent UMI barcode using recursive substitution error correction (RSEC). For targeted sequencing analysis, other UMI errors derived from library preparation steps or sequencing base deletions are later adjusted using distribution-based error correction (DBEC). **THUS DBEC is more stringent than RSEC.
## BD Handkbook page 53: Note: It is generally recommended to use <sample_name>_DBEC_MolsPerCell.csv for clustering analysis. Read counts for DBEC, read counts for RSEC, and molecule counts for RSEC are provided for reference. The RSEC files can be used when sequencing depth is so low that most genes do not pass the threshold for the DBEC molecular identifier adjustment algorithm to be applied; that is, low_depth in <sample_name>_UMI_Adjusted_Stats.csv.


## Per Rhapsody Manual: Note that targeted sequencing analysis uses RSEC and DBEC, while WTA sequencing analysis uses RSEC only. 
## Per Rhapsody Manual: Figure 6. Example results after applying RSEC and DBEC algorithms. For targeted sequencing analysis, if we consider only raw UMIs, the apparent total number of molecules continues to rise with sequencing depth, because the presence of sequencing and PCR errors contribute to unique UMIs. RSEC removes artifact molecules from single base substitutions in the UMI sequence. Further adjustment by DBEC removes artifact molecules originated from PCR errors. As a result, the number of molecules stabilizes with additional sequencing, indicating the library is sequenced to saturation.


```{r, warning=FALSE, message=FALSE, warning=FALSE, echo=F}
set.seed(1234)
#graphics.off()
#rm(list=ls())
#knitr::opts_chunk$set(fig.width=8, fig.height=6, dpi=300,echo=FALSE, warning=FALSE, message=FALSE)
#knitr::opts_chunk$set(dev.args=list(bg="transparent"))
#ROOT_DIR<-"~/Fred Hutchinson Cancer Research Center/GVHD - Hill Furlan Collaboration - General"
setwd("/fh/fast/hill_g/Motoko/Exp160/data")
ROOT_DIR<-"/fh/fast/hill_g/Motoko"
stem<-"Exp160"
DATA_DIR <- file.path(ROOT_DIR, stem, "data")      # SPECIFY HERE
RES_DIR  <- file.path(ROOT_DIR, stem, "res")     # SPECIFY HERE
RMD_DIR  <- file.path(ROOT_DIR, stem, "rmd")     # SPECIFY HERE
CDS_DIR <- file.path(ROOT_DIR, stem, "cds")
FIG_DIR <- file.path(ROOT_DIR,  stem, "figs")

ENR_DIR<-file.path(ROOT_DIR, "Bioinformatics Resources", "MSigDB")
SCRIPT_DIR<-file.path(ROOT_DIR, "develop", "SfScripts", "m3")

suppressPackageStartupMessages({
  library(Seurat)
  library(SeuratWrappers) # Needed to run monocle3; install as follows: devtools::install_github('satijalab/seurat-wrappers')
  library(scCustomize) # Needed to run SetUP.R (/fh/fast/hill_g/Albert/scSeq_Analyses/Setup Files)
    #Error: ERROR: no permission to install to directory ‘/home/ayeh/R/x86_64-pc-linux-gnu-library/4.3’
  library(SCpubr) 
  library(monocle3)
  library(m3addon) #plugin for monocle #options(repos = c(CRAN = "https://cloud.r-project.org"))
  library(ggplot2)
  library(ggplot2)
  library(patchwork)
  library(RColorBrewer)
  library(tidyverse)
  library(scCustomize)
  library(ArchR)
  library(viridis)
  library(ggpubr)
  library(Matrix)
  # devtools::install_github("immunogenomics/presto")
  
  #dyn.load('/app/software/ArrayFire/3.8.1/lib64/libaf.so.3')
  #library(RcppArrayFire)
  #library(viewmastR) # https://github.com/scfurl/viewmaster
})

#hg38<-getBSgenome("BSgenome.Hsapiens.UCSC.hg38" )

xfun::pkg_load2(c('base64enc', 'htmltools', 'mime'))
#install.packages("tidyverse")
#devtools::install_github("waltno/viewmaster")

SFtheme<-theme_bw(base_size=14) + 
  theme(panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_blank(), 
    panel.grid.major = element_blank(),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_rect(fill = "transparent"), 
    legend.key = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent",colour = NA))
theme_set(SFtheme)


clus_cols12 <- c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b", 
                 "#e377c2", "#7f7f7f", "#bcbd22", "#17becf", "#aec7e8", "#ffbb78")
clus_cols16 <- c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b", 
                 "#e377c2", "#7f7f7f", "#bcbd22", "#17becf", "#aec7e8", "#ffbb78", 
                 "#98df8a", "#ff9896", "#c5b0d5", "#c49c94")
clus_cols24 <- c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b", 
                 "#e377c2", "#7f7f7f", "#bcbd22", "#17becf", "#aec7e8", "#ffbb78", 
                 "#98df8a", "#ff9896", "#c5b0d5", "#c49c94", "#f7b6d2", "#c7c7c7", 
                 "#dbdb8d", "#9edae5", "#ad494a", "#8c6d31", "#e7ba52", "#17becf")
clus_cols36 <- c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b", 
                 "#e377c2", "#7f7f7f", "#bcbd22", "#17becf", "#aec7e8", "#ffbb78",
                 "#98df8a", "#ff9896", "#c5b0d5", "#c49c94", "#f7b6d2", "#c7c7c7", 
                 "#dbdb8d", "#9edae5", "#ad494a", "#8c6d31", "#e7ba52", "#17becf",
                 "#1f78b4", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#cab2d6",
                 "#6a3d9a", "#ff8c00", "#b15928", "#41ab5d", "#f03b20", "#807dba")

#pie(rep(1, length(col)), col = col)
datestamp<-"241108"

rna_cols<-paletteContinuous(n=8)[c(1:3, 6:8)]

```

```{r - loads RSEC data, echo=FALSE, include = TRUE}
## this uses RSEC
f<-list.files(file.path(DATA_DIR, paste0("cart", 1:2)), full.names = T)[grep("RSEC_MolsPerCell", list.files(file.path(DATA_DIR, paste0("cart", 1:2))))] #This line lists all files in two directories, cart1 and cart2, inside DATA_DIR.
files<-f[c(2,4)]

dat<-lapply(files, function(file){
t(as.matrix(data.table::fread(file), rownames=1))}) # For each file in files, it reads the file into a matrix using data.table::fread. The t() function transposes each matrix, making rows into columns and columns into rows. dat will contain two transposed matrices from the selected files.

rn<-intersect(rownames(dat[[1]]), rownames(dat[[2]])) # rn is a list of row names common to both matrices.
rn<-sapply(strsplit(rn, "\\|"), "[[", 1)  # The strsplit(rn, "\\|") splits each row name by the "|" delimiter. sapply(..., "[[", 1) extracts the first part before the delimiter for each name.

dat2<-lapply(dat, function(mat){
  rownames(mat)<-sapply(strsplit(rownames(mat), "\\|"), "[[", 1)
  mat[rn,]
}) #This modifies the row names of each matrix by keeping only the first part (before the "|"). Each matrix is then subsetted to include only the rows that match the names in rn, creating a new list dat2 with the processed matrices.

colnames(dat2[[1]])<-paste0("cart1_", colnames(dat2[[1]]))
colnames(dat2[[2]])<-paste0("cart2_", colnames(dat2[[2]]))
mat<-Matrix(do.call(cbind, dat2), sparse=T) # The two matrices in dat2 are combined column-wise (i.e., bound together as columns) using do.call(cbind, dat2). The Matrix() function with sparse=T creates a sparse matrix from the combined data, saving memory by storing only non-zero entries.
colnames(mat)<-make.unique(colnames(mat)) # This line ensures that all column names in the combined matrix mat are unique by appending numbers where duplicate names are found.

rmat<-mat[5:dim(mat)[1],] # rmat contains the main GEX data starting from the 8th row.
pmat<-mat[1:4,] # pmat contains the first 7 rows and holds the ADT data

seu<-CreateSeuratObject(counts = rmat)
seu$ADT<-CreateAssayObject(counts = pmat[,match(colnames(pmat), Cells(seu))])

#DefaultAssay(seu)
## ADT Features:
# [1] "CD326"        
# [5] "CD45.2"             "H-2Kb"              "I-A-I-E"  

seu <- NormalizeData(seu, normalization.method = "LogNormalize", scale.factor = 10000)
seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 3000)
seu <- ScaleData(seu)
seu <- RunPCA(seu, features = VariableFeatures(object = seu), npcs = 50)
#ElbowPlot(seu, 50)
seu <- FindNeighbors(seu, dims = 1:40)
seu <- FindClusters(seu, resolution = 0.5)
seu <- RunUMAP(seu, dims = 1:40)

DimPlot(seu)
seu$cart<-strsplit(Cells(seu), "_") %>% sapply("[[", 1)
seu[["percent.mt"]] <- PercentageFeatureSet(seu, pattern = "^mt-")
seu$logUMI<-log10(seu$nCount_RNA)
seu$log_RNA<-log10(seu$nCount_RNA)


## For QC presentation ##
VlnPlot(seu, features = c("nFeature_RNA", "log_RNA", "logUMI", "percent.mt"), group.by = "cart", ncol = 4, pt.size = 0)
DimPlot(seu, group.by = "cart")
FeaturePlot_scCustom(seu, features = "percent.mt")
####


## Now filter by RNA count
DefaultAssay(seu_sub)<-"RNA"

#seu_sub <- subset(seu, subset = log_RNA > 3.25 & log_RNA < 4.25 & percent.mt < 40) 
seu_sub <- subset(seu, subset = nCount_RNA > 1000 & nCount_RNA < 50000 & percent.mt < 40) # This gives us 8,032 samples after filtering (from 15,288 before0)
seu_sub <- NormalizeData(seu_sub, normalization.method = "LogNormalize", scale.factor = 10000)
seu_sub <- FindVariableFeatures(seu_sub, selection.method = "vst", nfeatures = 3000)
seu_sub <- ScaleData(seu_sub)
seu_sub <- RunPCA(seu_sub, features = VariableFeatures(object = seu_sub), npcs = 50)
#ElbowPlot(seu_sub, 50)
seu_sub <- FindNeighbors(seu_sub, dims = 1:40)
seu_sub <- FindClusters(seu_sub, resolution = 0.15)
seu_sub <- RunUMAP(seu_sub, dims = 1:40)

## Data output for QC ##
DimPlot(seu_sub)
VlnPlot(seu_sub, features = c("nFeature_RNA", "log_RNA", "logUMI", "percent.mt"), group.by = "cart", ncol = 4, pt.size = 0)
DimPlot(seu_sub, group.by = "cart")
FeaturePlot_scCustom(seu_sub, features = "percent.mt")
#####

## Normalize ADT Data ##
DefaultAssay(seu_sub)<-"ADT"
seu_sub <- NormalizeData(seu_sub, normalization.method = "CLR")
seu_sub <- ScaleData(seu_sub)

DimPlot(seu_sub, cols = clus_cols36)

#### Temporary Save Point Here ####
#The RDS file below has removed multiplets / those without proper hashtags, as well as filter by RNA counts and mito genes.  
#Also, 4 groups (female IFNy +/- and male IFNy +/-) have been added
saveRDS(seu_sub, file.path(CDS_DIR, "241108v2_processed.RDS"))
####
```


```{r - analysis of clusters, echo=FALSE, include=TRUE}
seu<-readRDS(file.path(CDS_DIR, "241108v2_processed.RDS"))

pdf(file.path(FIG_DIR, paste0("[Final] - Initial Clustering, res = 0.1", ".pdf")), width=6, height=6)
DimPlot(seu, cols = clus_cols36)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Initial Clustering by cart", ".pdf")), width=6, height=6)
DimPlot(seu,  group.by = "cart")
dev.off()


FeaturePlot_scCustom(seu, features = "Mki67")
FeaturePlot_scCustom(seu, features = "percent.mt")

DefaultAssay(seu)<-"ADT"
pdf(file.path(FIG_DIR, paste0("[Final] - I-A-I-E", ".pdf")), width=6, height=6)
FeaturePlot_scCustom(seu, features = "I-A-I-E", max.cutoff = "q99")
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - H-2Kb", ".pdf")), width=6, height=6)
FeaturePlot_scCustom(seu, features = "H-2Kb", max.cutoff = "q99")
dev.off()




FeaturePlot_scCustom(seu, features = "Lgr5", max.cutoff = "q99")

FeaturePlot_scCustom(seu, features = "Sis", max.cutoff = "q99") #Sucrose isomaltase
FeaturePlot_scCustom(seu, features = "Tgfbi", max.cutoff = "q99")
FeaturePlot_scCustom(seu, features = "Olfm4", max.cutoff = "q99")

FeaturePlot_scCustom(seu, features = "Oat", max.cutoff = "q99")
FeaturePlot_scCustom(seu, features = "Ppp1r1b", max.cutoff = "q99")
FeaturePlot_scCustom(seu, features = "Apol10a", max.cutoff = "q99")


DefaultAssay(seu)<-"RNA"


markers_1_vs_3 <- FindMarkers(seu, ident.1 = 1, ident.2 = 3) #Helped identify Chgb and Cpe = enteroendocrine cells

write.csv(markers_1_vs_3,"1 vs 3.csv")
markers_0_vs_3 <- FindMarkers(seu, ident.1 = 0, ident.2 = 3) #Helped identify Chgb and Cpe = enteroendocrine cells
write.csv(markers_0_vs_3,"0 vs 3.csv")


markers_1_vs_5 <- FindMarkers(seu, ident.1 = 1, ident.2 = 5) #Helped identify Chgb and Cpe = enteroendocrine cells

pdf(file.path(FIG_DIR, paste0("[Final] - Chromogranin B", ".pdf")), width=6, height=6)
FeaturePlot_scCustom(seu, features = "Chgb")
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Muc2", ".pdf")), width=6, height=6)
FeaturePlot_scCustom(seu, features = "Muc2") # Goblet/Paneth
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Trpm5", ".pdf")), width=6, height=6)
FeaturePlot_scCustom(seu, features = "Trpm5") # Tufts
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Lyz1", ".pdf")), width=6, height=6)
FeaturePlot_scCustom(seu, features = "Lyz1") # Lysozyme (Paneth)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Mmp7", ".pdf")), width=6, height=6)
FeaturePlot_scCustom(seu, features = "Mmp7") # Mmp7 (Paneth)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Pou2f3", ".pdf")), width=6, height=6)
FeaturePlot_scCustom(seu, features = "Pou2f3") # Pou2f3 (Tufts)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Anpep", ".pdf")), width=6, height=6)
FeaturePlot_scCustom(seu, features = "Anpep") # Aminopeptidase N (Enterocyte)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Epcam", ".pdf")), width=6, height=6)
FeaturePlot_scCustom(seu, features = "Epcam") # EPCAM
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Lgr5", ".pdf")), width=6, height=6)
FeaturePlot_scCustom(seu, features = "Lgr5") # ISC
dev.off()


FeaturePlot_scCustom(seu, features = "Bmi1") # ISC
FeaturePlot_scCustom(seu, features = "Ephb2") # ISC
markers_1_vs_4 <- FindMarkers(seu, ident.1 = 1, ident.2 = 4) #Helped identify Chgb and Cpe = enteroendocrine cells
FeaturePlot_scCustom(seu, features = "Ccl25") # EPCAM

pdf(file.path(FIG_DIR, paste0("[Final] - Markers", ".pdf")), width=6, height=6)
DotPlot(seu, features = c("Chgb", "Cpe", "Muc2", "Lyz1", "Mmp7", "Trpm5", "Pou2f3", "Anpep", "Epcam","Lgr5","Ccl25")) + RotatedAxis()
dev.off()

```

```{r - expression markers, echo=FALSE, include=TRUE}


# [1] "CD326"        
# [5] "CD45.2"             "H-2Kb"              "I-A-I-E"  

DefaultAssay(seu)<-"ADT"

my_comparisons <- list( c("0","1"),c("0","2"),c("0","3"),c("0","4"),c("0","5")) 

pdf(file.path(FIG_DIR, paste0("[Final] - I-A-I-E", ".pdf")), width=6, height=6)
FeaturePlot_scCustom(seu, features = "I-A-I-E") # Tufts
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - I-A-I-E (violin)" , ".pdf")), width=6, height=6)
VlnPlot(seu, features = c("I-A-I-E"), group.by="seurat_clusters", col = clus_cols36,  pt.size = 0) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + stat_compare_means(comparisons = my_comparisons, method = "wilcox.test") + ylim(0, 10)
dev.off()


pdf(file.path(FIG_DIR, paste0("[Final] - H-2Kb", ".pdf")), width=6, height=6)
FeaturePlot_scCustom(seu, features = "H-2Kb") # Tufts
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - H-2Kb (violin)" , ".pdf")), width=6, height=6)
VlnPlot(seu, features = c("H-2Kb"), group.by="seurat_clusters", col = clus_cols36, pt.size = 0) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + stat_compare_means(comparisons = my_comparisons, method = "wilcox.test") + ylim(0, 12)
dev.off()


```

```{r - GSEA, echo=FALSE, include=FALSE}

#========== Figure 5E GSEA ==========#
#---- GSEA analysis using fsgea ----#
#https://crazyhottommy.github.io/scRNA-seq-workshop-Fall-2019/scRNAseq_workshop_3.html
library(presto)
library(dplyr)
library(ggplot2)
library(tibble)
library(fgsea)
library(msigdbr)

seu.genes <- wilcoxauc(seu, 'seurat_clusters')

# for GSEA, we need the information of all genes, Seurat is just too slow if we test
# all 20,000 genes. instead let's try presto which performs a fast Wilcoxon rank sum test 

#---- Hallmark Gene Sets ----#
m_t2g <- msigdbr(species = "Mus musculus", category = "H") %>% dplyr::select(gs_name, gene_symbol) 
fgsea_sets<- m_t2g %>% split(x = .$gene_symbol, f = .$gs_name) #Collects names of the genes of interest

#----- Comparing gene sets -----#
pdf(file.path(FIG_DIR, paste0("[Final] - GOCC_MHC_CLASS_II_PROTEIN_COMPLEX", ".pdf")), width=6, height=6)
bac<-m_t2g[m_t2g$gs_name == "GOCC_MHC_CLASS_II_PROTEIN_COMPLEX",]$gene_symbol
DefaultAssay(seu)<-"RNA"
seu<-AddModuleScore(seu, features = list(bac), name = "MHC_CLASS_II_PROTEIN_COMPLEX")
FeaturePlot_scCustom(seu, features = "MHC_CLASS_II_PROTEIN_COMPLEX1", label = F, label.size = 2, pt.size = 1, max.cutoff = "q99")+scale_color_gradientn(colors = rna_cols)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - GOCC_MHC_CLASS_II_PROTEIN_COMPLEX violin", ".pdf")), width=6, height=6)
VlnPlot(seu, features = c("MHC_CLASS_II_PROTEIN_COMPLEX1"), group.by="seurat_clusters", pt.size = 0) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + stat_compare_means(comparisons = my_comparisons, method = "wilcox.test") + ylim(0, 5)
dev.off()


pdf(file.path(FIG_DIR, paste0("[Final] - GOCC_MHC_CLASS_I_PROTEIN_COMPLEX", ".pdf")), width=6, height=6)
bac<-m_t2g[m_t2g$gs_name == "GOCC_MHC_CLASS_I_PROTEIN_COMPLEX",]$gene_symbol
DefaultAssay(seu)<-"RNA"
seu<-AddModuleScore(seu, features = list(bac), name = "GOCC_MHC_CLASS_I_PROTEIN_COMPLEX")
FeaturePlot_scCustom(seu, features = "GOCC_MHC_CLASS_I_PROTEIN_COMPLEX1", label = F, label.size = 2, pt.size = 1, max.cutoff = "q99")+scale_color_gradientn(colors = rna_cols)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - GOCC_MHC_CLASS_I_PROTEIN_COMPLEX violin", ".pdf")), width=6, height=6)
VlnPlot(seu, features = c("GOCC_MHC_CLASS_I_PROTEIN_COMPLEX1"), group.by="seurat_clusters", pt.size = 0) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + stat_compare_means(comparisons = my_comparisons, method = "wilcox.test") + ylim(0, 2.5)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - GOBP_ANTIGEN_PROCESSING_AND_PRESENTATION", ".pdf")), width=6, height=6)
bac<-m_t2g[m_t2g$gs_name == "GOBP_ANTIGEN_PROCESSING_AND_PRESENTATION",]$gene_symbol
DefaultAssay(seu)<-"RNA"
seu<-AddModuleScore(seu, features = list(bac), name = "GOBP_ANTIGEN_PROCESSING_AND_PRESENTATION")
FeaturePlot_scCustom(seu, features = "GOBP_ANTIGEN_PROCESSING_AND_PRESENTATION1", label = F, label.size = 2, pt.size = 1, max.cutoff = "q99")+scale_color_gradientn(colors = rna_cols)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - GOBP_ANTIGEN_PROCESSING_AND_PRESENTATION violin", ".pdf")), width=6, height=6)
VlnPlot(seu, features = c("GOBP_ANTIGEN_PROCESSING_AND_PRESENTATION1"), group.by="seurat_clusters", pt.size = 0) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + stat_compare_means(comparisons = my_comparisons, method = "wilcox.test") + ylim(0, 1)
dev.off()


```

```{r - custom gene sets, echo=FALSE, include=FALSE}

#Gene sets:
#Grun D et al Nature 2015 (PMID: 26287467)
#Haber AL et al Nature 2017 (PMID: 29144463)
#Biton M et at. Cell 2018 (PMID: 30392957)
#Hadjsombati MS et al. (Tuft cell) Immunity 2018 (PMID: 30021144)
#Wang et al. JEM 2019 (PMID: 31753849)
#Zhao Y et al. Mucosal Immunol 2020 (PMID: 31900405)


#Stem cell: 	Lgr5, Olfm4, Slc12a2, Ascl2, Axin2, Gkn3
#Transit-amplifying cells (TA):	 Olfm4, Scl12a2, (nusap1 (found in human))
#Goblet:	Muc2, Agr2, Clca3, Tff3, Ang4, Defa31
#Tuft:		Dclk1, Trpm5, Gnat3 (α-gustducin), CD24a, Krt8, Krt18, Gfi1b, Il25, Ptgs1, Alox5, Ptprc, Chat , Rac2 (specific)
#Paneth:	Mptx1 (specific), Ang4, Defa31, Defa17, Defa24, Defa22, Lyz1, Itln1 (Paneth in mouse)
#Enteroendocrine: Reg4 (specific), Chga, Chgb, Tph1, Tac1, Neuro3, 
#Enterocytes: Alpi, Apoa1, Apoa4, Fabp1, 


custom_gene_sets <- list(
  "Stem_Cell" = c("Lgr5", "Olfm4", "Slc12a2", "Ascl2", "Axin2", "Gkn3"),
  "Transit_Amplifying_Cells" = c("Olfm4", "Scl12a2"), # nusap1 found in human
  "Goblet" = c("Muc2", "Agr2", "Clca3", "Tff3", "Ang4", "Defa31"),
  "Tuft" = c("Dclk1", "Trpm5", "Gnat3", "CD24a", "Krt8", "Krt18", "Gfi1b", "Il25", "Ptgs1", "Alox5", "Ptprc", "Chat", "Rac2"), # Gnat3 is α-gustducin, Rac2 specific
  "Paneth" = c("Mptx1", "Ang4", "Defa31", "Defa17", "Defa24", "Defa22", "Lyz1", "Itln1"), # Mptx1 specific, Itln1 Paneth in mouse
  "Enteroendocrine" = c("Reg4", "Chga", "Chgb", "Tph1", "Tac1", "Neuro3"), # Reg4 specific
  "Enterocytes" = c("Alpi", "Apoa1", "Apoa4", "Fabp1")
)

pdf(file.path(FIG_DIR, paste0("[Final] - Custom Stem_Cell", ".pdf")), width=6, height=6)
bac<-custom_gene_sets$Stem_Cell
DefaultAssay(seu)<-"RNA"
seu<-AddModuleScore(seu, features = list(bac), name = "Stem_Cell")
FeaturePlot_scCustom(seu, features = "Stem_Cell1", label = F, label.size = 1, pt.size = 1, max.cutoff = "q99")+scale_color_gradientn(colors = rna_cols)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Custom Transit_Amplifying_Cells", ".pdf")), width=6, height=6)
bac<-custom_gene_sets$Transit_Amplifying_Cells
DefaultAssay(seu)<-"RNA"
seu<-AddModuleScore(seu, features = list(bac), name = "Transit_Amplifying_Cells")
FeaturePlot_scCustom(seu, features = "Transit_Amplifying_Cells1", label = F, label.size = 1, pt.size = 1, max.cutoff = "q99")+scale_color_gradientn(colors = rna_cols)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Custom Goblet", ".pdf")), width=6, height=6)
bac<-custom_gene_sets$Goblet
DefaultAssay(seu)<-"RNA"
seu<-AddModuleScore(seu, features = list(bac), name = "Goblet")
FeaturePlot_scCustom(seu, features = "Goblet1", label = F, label.size = 1, pt.size = 1, max.cutoff = "q99")+scale_color_gradientn(colors = rna_cols)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Custom Tuft", ".pdf")), width=6, height=6)
bac<-custom_gene_sets$Tuft
DefaultAssay(seu)<-"RNA"
seu<-AddModuleScore(seu, features = list(bac), name = "Tuft")
FeaturePlot_scCustom(seu, features = "Tuft1", label = F, label.size = 1, pt.size = 1, max.cutoff = "q99")+scale_color_gradientn(colors = rna_cols)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Custom Paneth", ".pdf")), width=6, height=6)
bac<-custom_gene_sets$Paneth
DefaultAssay(seu)<-"RNA"
seu<-AddModuleScore(seu, features = list(bac), name = "Paneth")
FeaturePlot_scCustom(seu, features = "Paneth1", label = F, label.size = 1, pt.size = 1, max.cutoff = "q99")+scale_color_gradientn(colors = rna_cols)
dev.off()

pdf(file.path(FIG_DIR, paste0("[Final] - Custom Enteroendocrine", ".pdf")), width=6, height=6)
bac<-custom_gene_sets$Enteroendocrine
DefaultAssay(seu)<-"RNA"
seu<-AddModuleScore(seu, features = list(bac), name = "Enteroendocrine")
FeaturePlot_scCustom(seu, features = "Enteroendocrine1", label = F, label.size = 1, pt.size = 1, max.cutoff = "q99")+scale_color_gradientn(colors = rna_cols)
dev.off()


pdf(file.path(FIG_DIR, paste0("[Final] - Custom Enterocytes", ".pdf")), width=6, height=6)
bac<-custom_gene_sets$Enterocytes
DefaultAssay(seu)<-"RNA"
seu<-AddModuleScore(seu, features = list(bac), name = "Enterocytes")
FeaturePlot_scCustom(seu, features = "Enterocytes1", label = F, label.size = 1, pt.size = 1, max.cutoff = "q99")+scale_color_gradientn(colors = rna_cols)
dev.off()

```

# Pseudotime Trajectory
```{r pseudotime}
cds<-seurat_to_monocle3(seu)
#troubleshooting
cds@int_colData@listData[["reducedDims"]]@listData<-cds@int_colData@listData[["reducedDims"]]@listData[1]

plot_cells(cds, genes = c("Sell", "Lef1"), scale_to_range = F)
cds<-cluster_cells(cds, resolution = 0.001)
plot_cells(cds)

cds@clusters@listData[["UMAP"]][["clusters"]]<-clusters(cds)
cds<-learn_graph(cds, use_partition = T, close_loop = T)
cds<- order_cells(cds)
plot_cells(cds, color_cells_by = "pseudotime", show_trajectory_graph = T)

seu$pseudotime<-pseudotime(cds)
pdf("../figs/cd4_pseudotime.pdf", width = 5, height = 4)
FeaturePlot(seu, features = "pseudotime")+scale_color_gradientn(colors = viridis::plasma(n=10))&NoAxes()
dev.off()
```