---
title: "Exp160 - Figure 4 and Figure S6"
author: "Koyama et al., Science Immunology"
output: html_document
---

# Exp160 — Figure 4 and Figure S6
#Ileal intestinal epithelial cells from **cartridge1** (naive WT B6) and **cartridge2** (24 h post-TBI WT B6).

```{r setup}
suppressPackageStartupMessages({
  library(Seurat); library(ggplot2); library(dplyr); library(patchwork)
  library(scCustomize); library(ggpubr); library(Matrix)
})
```

## Setting up
**Edit one line — `DATA`.** Download this experiment's processed matrices
from NCBI GEO [GSE348009](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE348009)
to wherever you like and point `DATA` at that folder. It must contain
`cartridge1/` and `cartridge2/`.

Output is written to
`Exp160_fig4_figS6/` in the working directory.

```{r paths}
## ==========  EDIT THIS ONE LINE  ==========
DATA <- "~/GSE348009/Exp160"
## ==========================================
```

## Helper functions
Loading and processing used below.
```{r functions}
suppressPackageStartupMessages({
  library(Matrix)
  library(data.table)
  library(Seurat)
})

## Path to the filtered per-cell molecule matrix for one cartridge.
## `which` is "RSEC" or "DBEC".
bd_matrix_path <- function(dir, which = c("RSEC", "DBEC")) {
  f <- list.files(dir, full.names = TRUE)
  hit <- grep(paste0("_", which, "_MolsPerCell\\.csv$"), f, value = TRUE)
  hit
}

## Number of AbSeq targets, counted from the file's own header line.
bd_n_pabo <- function(path) {
  con <- file(path, "r"); on.exit(close(con))
  repeat {
    line <- readLines(con, n = 1L)
    if (!startsWith(line, "#")) break
  }
  cols <- strsplit(line, ",", fixed = TRUE)[[1]]
  sum(endsWith(cols, "|pAbO"))
}

## Read one or more cartridges into a single sparse matrix.
##   dirs      named character vector: name = cart prefix, value = directory
##   n_pabo    expected number of AbSeq rows (checked, not assumed)
bd_load <- function(dirs, which = "RSEC", n_pabo = NULL) {
  paths <- vapply(dirs, bd_matrix_path, character(1), which = which)
  k <- unique(vapply(paths, bd_n_pabo, integer(1)))
  message(sprintf("  %s: %d cartridge(s), %d AbSeq targets", which, length(paths), k))

  dat <- lapply(paths, function(p) {
    message("  reading ", basename(p))
    t(as.matrix(data.table::fread(p), rownames = 1))
  })

  rn <- Reduce(intersect, lapply(dat, rownames))
  rn <- vapply(strsplit(rn, "\\|"), `[[`, character(1), 1L)
  dat <- lapply(dat, function(m) {
    rownames(m) <- vapply(strsplit(rownames(m), "\\|"), `[[`, character(1), 1L)
    m[rn, , drop = FALSE]
  })
  for (i in seq_along(dat))
    colnames(dat[[i]]) <- paste0(names(dirs)[i], "_", colnames(dat[[i]]))

  mat <- Matrix(do.call(cbind, dat), sparse = TRUE)
  colnames(mat) <- make.unique(colnames(mat))
  list(rna = mat[(k + 1L):nrow(mat), , drop = FALSE],
       adt = mat[1L:k, , drop = FALSE],
       n_pabo = k)
}

## Build the Seurat object: RNA counts, the ADT (protein) assay, a cartridge
## label per cell, mitochondrial percentage and log10 UMI counts.
bd_seurat <- function(loaded) {
  seu <- CreateSeuratObject(counts = loaded$rna)
  seu[["ADT"]] <- CreateAssayObject(
    counts = loaded$adt[, match(colnames(loaded$adt), Cells(seu)), drop = FALSE])
  seu$cart <- vapply(strsplit(Cells(seu), "_"), `[[`, character(1), 1L)
  seu[["percent.mt"]] <- PercentageFeatureSet(seu, pattern = "^mt-")
  seu$logUMI  <- log10(seu$nCount_RNA)
  seu$log_RNA <- log10(seu$nCount_RNA)
  seu
}

## LogNormalize -> 3000 variable features -> scale -> PCA(50) -> neighbours,
## clusters, UMAP.
bd_process <- function(seu, dims, resolution, seed = 1234) {
  set.seed(seed)
  DefaultAssay(seu) <- "RNA"
  seu <- NormalizeData(seu, normalization.method = "LogNormalize", scale.factor = 10000)
  seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 3000)
  seu <- ScaleData(seu)
  # Seurat's per-function seed defaults are left as they are: FindClusters
  # (random.seed = 0), RunPCA/RunUMAP (seed.use = 42) and AddModuleScore
  # (seed = 1). Overriding them is still deterministic but gives 8 clusters
  # where Figure 4 has 7.
  seu <- RunPCA(seu, features = VariableFeatures(object = seu), npcs = 50)
  seu <- FindNeighbors(seu, dims = dims)
  seu <- FindClusters(seu, resolution = resolution)
  seu <- RunUMAP(seu, dims = dims)
  seu
}

## CLR normalisation of the protein assay.
bd_process_adt <- function(seu) {
  DefaultAssay(seu) <- "ADT"
  seu <- NormalizeData(seu, normalization.method = "CLR")
  seu <- ScaleData(seu)
  DefaultAssay(seu) <- "RNA"
  seu
}

## The continuous expression palette: ArchR::paletteContinuous(n=8)[c(1:3,6:8)].
bd_rna_cols <- function() {
  if (!requireNamespace("ArchR", quietly = TRUE))
    stop("ArchR is not installed; it defines rna_cols (paletteContinuous). ",
         "Run common/scripts/install_r_deps.sh first - do not substitute another palette.")
  ArchR::paletteContinuous(n = 8)[c(1:3, 6:8)]
}

bd_clus_cols36 <- c(
  "#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b",
  "#e377c2", "#7f7f7f", "#bcbd22", "#17becf", "#aec7e8", "#ffbb78",
  "#98df8a", "#ff9896", "#c5b0d5", "#c49c94", "#f7b6d2", "#c7c7c7",
  "#dbdb8d", "#9edae5", "#ad494a", "#8c6d31", "#e7ba52", "#17becf",
  "#1f78b4", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#cab2d6",
  "#6a3d9a", "#ff8c00", "#b15928", "#41ab5d", "#f03b20", "#807dba")

bd_theme <- function() {
  library(ggplot2)
  theme_bw(base_size = 14) +
    theme(panel.background   = element_rect(fill = "transparent", colour = NA),
          panel.grid.minor   = element_blank(),
          panel.grid.major   = element_blank(),
          legend.background  = element_rect(fill = "transparent"),
          legend.box.background = element_rect(fill = "transparent"),
          legend.key         = element_rect(fill = "transparent", colour = NA),
          plot.background    = element_rect(fill = "transparent", colour = NA))
}

## Write the R and package versions beside the results.
bd_session <- function(path) {
  writeLines(c(capture.output(sessionInfo()), "",
               paste("run at", format(Sys.time(), "%Y-%m-%d %H:%M:%S"))), path)
}
```

```{r output-paths}
OUT <- "Exp160_fig4_figS6"
FIG <- file.path(OUT, "figs"); CDS <- file.path(OUT, "cds")
for (d in c(FIG, CDS)) dir.create(d, recursive = TRUE, showWarnings = FALSE)

theme_set(bd_theme())
rna_cols    <- bd_rna_cols()
clus_cols36 <- bd_clus_cols36
SEED <- 1234
pdfout <- function(name, w = 6, h = 6) pdf(file.path(FIG, paste0(name, ".pdf")), width = w, height = h)

cat("== Exp160 Fig 4 / S6 ==\n")
```

## 1. load cart1 + cart2

```{r sec1-load-cart1-cart2}
dirs <- c(cart1 = file.path(DATA, "cartridge1"),
          cart2 = file.path(DATA, "cartridge2"))
loaded <- bd_load(dirs, which = "RSEC", n_pabo = 4)
seu <- bd_seurat(loaded)
cat("   cells before filtering:", ncol(seu), "\n")
```

## 2. QC filter, then re-process
```{r sec2-qc-filter-then-re-process}

seu <- bd_process(seu, dims = 1:40, resolution = 0.5, seed = SEED)
cat("   first pass:", ncol(seu), "cells,",
    nlevels(seu$seurat_clusters), "clusters at resolution 0.5\n")
seu_sub <- subset(seu, subset = nCount_RNA > 1000 & nCount_RNA < 50000 & percent.mt < 40)
cat("   cells after filtering :", ncol(seu_sub))
seu_sub <- bd_process(seu_sub, dims = 1:40, resolution = 0.15, seed = SEED)
seu_sub <- bd_process_adt(seu_sub)
saveRDS(seu_sub, file.path(CDS, "cart1_cart2_processed.RDS"))
seu <- seu_sub

## QC panels (not published, kept as the run's own evidence)
pdfout("QC - violin by cartridge", 10, 4)
print(VlnPlot(seu, features = c("nFeature_RNA","log_RNA","logUMI","percent.mt"),
              group.by = "cart", ncol = 4, pt.size = 0))
dev.off()
```

## 3. Figure 4B

```{r sec3-figure-4b}
pdfout("[Fig4B] Initial Clustering, res=0.15, dim=40")
print(DimPlot(seu, cols = clus_cols36)); dev.off()
pdfout("[Fig4B] Initial Clustering by cart")
print(DimPlot(seu, group.by = "cart")); dev.off()

```

## 4. Figure 4C - MSigDB C5 module scores

```{r sec4-figure-4c-msigdb-c5-module-scores}}
options(warn = -1)
suppressPackageStartupMessages({ library(msigdbr) })
m_t2g <- msigdbr(species = "Mus musculus", category = "C5") %>%
  dplyr::select(gs_name, gene_symbol)

my_comparisons <- list(c("0","1"),c("0","2"),c("0","3"),c("0","4"),c("0","5"),c("0","6"))

gs_panels <- list(
  list(set = "GOCC_MHC_CLASS_II_PROTEIN_COMPLEX", name = "MHC_CLASS_II_PROTEIN_COMPLEX", ymax = 5),
  list(set = "GOCC_MHC_CLASS_I_PROTEIN_COMPLEX",  name = "MHC_CLASS_I_PROTEIN_COMPLEX",  ymax = 2.5),
  list(set = "GOBP_ANTIGEN_PROCESSING_AND_PRESENTATION",
       name = "ANTIGEN_PROCESSING_AND_PRESENTATION", ymax = 1))

DefaultAssay(seu) <- "RNA"
for (p in gs_panels) {
  genes <- m_t2g[m_t2g$gs_name == p$set, ]$gene_symbol
  stopifnot(length(genes) > 0)
  set.seed(SEED)
  seu <- AddModuleScore(seu, features = list(genes), name = p$name)
  feat <- paste0(p$name, "1")
  pdfout(paste0("[Fig4C] ", p$set))
  print(FeaturePlot_scCustom(seu, features = feat, label = FALSE, pt.size = 1,
                             max.cutoff = "q99") + scale_color_gradientn(colors = rna_cols))
  dev.off()
  pdfout(paste0("[Fig4C] ", p$set, " violin"))
  print(VlnPlot(seu, features = feat, group.by = "seurat_clusters",
                col = clus_cols36, pt.size = 0) +
        stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
        stat_compare_means(comparisons = my_comparisons, method = "wilcox.test") +
        ylim(0, p$ymax))
  dev.off()
}
```

## 5. Figure 4D / S6B - surface protein

```{r sec5-figure-4d-surface-protein}
options(warn = -1)
DefaultAssay(seu) <- "ADT"
for (ab in list(list(f = "I-A-I-E", ymax = 12), list(f = "H-2Kb", ymax = 14))) {
  pdfout(paste0("[Fig4D] ", ab$f))
  print(FeaturePlot_scCustom(seu, features = ab$f, max.cutoff = "q99")); dev.off()
  pdfout(paste0("[Fig4D] ", ab$f, " (violin)"))
  print(VlnPlot(seu, features = ab$f, group.by = "seurat_clusters",
                col = clus_cols36, pt.size = 0) +
        stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
        stat_compare_means(comparisons = my_comparisons, method = "wilcox.test") +
        ylim(0, ab$ymax))
  dev.off()
}
```

## 6. Figure S6A - published cell-subset signatures

Gene sets, from:
  Grun 2015 PMID 26287467; Haber 2017 PMID 29144463; Biton 2018 PMID 30392957;
  Hadjsombati 2018 PMID 30021144; Wang 2019 PMID 31753849; Zhao 2020 PMID 31900405;
  Yin 2023 doi 10.1038/s41421-023-00578-4; Malagola 2024 (Atad2/Stmn1/Smarca5/Dek)
```{r sec6-figure-s6a-published-cell-subset-signatures}
options(warn = -1)
custom_gene_sets <- list(
  Stem_Cell                = c("Lgr5","Olfm4","Slc12a2","Ascl2","Axin2","Gkn3"),
  Goblet                   = c("Muc2","Agr2","Clca3","Tff3","Ang4","Defa31"),
  Tuft                     = c("Dclk1","Trpm5","Gnat3","CD24a","Krt8","Krt18","Gfi1b",
                               "Il25","Ptgs1","Alox5","Ptprc","Chat","Rac2"),
  Paneth                   = c("Mptx1","Ang4","Defa31","Defa17","Defa24","Defa22","Lyz1","Itln1"),
  Enteroendocrine          = c("Reg4","Chga","Chgb","Tph1","Tac1","Neuro3"),
  Enterocytes              = c("Alpi","Apoa1","Apoa4","Fabp1"),
  Partial_IEC              = c("Top2a","Apol10a","Car4","Ube2c"),
  Malagola_2024            = c("Atad2","Stmn1","Smarca5","Dek"))

DefaultAssay(seu) <- "RNA"
for (nm in names(custom_gene_sets)) {
  set.seed(SEED)
  seu <- AddModuleScore(seu, features = list(custom_gene_sets[[nm]]), name = nm)
  pdfout(paste0("[FigS6A] Custom ", nm))
  print(FeaturePlot_scCustom(seu, features = paste0(nm, "1"), label = FALSE,
                             pt.size = 1, max.cutoff = "q99") +
        scale_color_gradientn(colors = rna_cols))
  dev.off()
}
```

## 7. Figure S6C - ISC / progenitor markers

```{r sec7-figure-s6c-isc-progenitor-markers}
for (g in c("Lgr5","Lgr4","Fgfbp1")) {
  pdfout(paste0("[FigS6C] ", g)); print(FeaturePlot_scCustom(seu, features = g)); dev.off()
}
```

## 8. Figure S6D/E - dot plots

```{r sec8-figure-s6d-e-dot-plots}
DefaultAssay(seu) <- "RNA"
pdfout("[FigS6D] Markers GSEA dotplot", 4, 6)
print(DotPlot(seu, features = c("MHC_CLASS_II_PROTEIN_COMPLEX1",
                                "MHC_CLASS_I_PROTEIN_COMPLEX1",
                                "ANTIGEN_PROCESSING_AND_PRESENTATION1")) +
      scale_x_discrete(labels = function(x) substr(x, 1, 15)) + RotatedAxis())
dev.off()
DefaultAssay(seu) <- "ADT"
pdfout("[FigS6E] Markers ADT dotplot", 4, 6)
print(DotPlot(seu, features = c("I-A-I-E", "H-2Kb")) + RotatedAxis()); dev.off()
DefaultAssay(seu) <- "RNA"
```

## 9. Figure S6B - Monocle3 pseudotime

The pseudotime root is the stem-cell cluster, as stated in the Figure S6B
legend. It is identified as the cluster with the highest mean Stem_Cell module
score, and the root is the principal-graph vertex closest to that cluster's
UMAP centroid.

```{r sec9-figure-s6b-monocle3-pseudotime}
suppressPackageStartupMessages({ library(monocle3); library(SeuratWrappers) })
set.seed(SEED)
cds <- as.cell_data_set(seu)
cds <- cluster_cells(cds, resolution = 1e-3)   # monocle3's default random_seed
cds <- learn_graph(cds, use_partition = TRUE, close_loop = TRUE)
cds@clusters@listData[["UMAP"]][["clusters"]] <- clusters(cds)

stem_score  <- tapply(seu$Stem_Cell1, seu$seurat_clusters, mean)
stem_clust  <- names(which.max(stem_score))
cat("   pseudotime root cluster (highest Stem_Cell score):", stem_clust,
    sprintf("(score %.3f)\n", max(stem_score)))
centroid <- colMeans(Embeddings(seu, "umap")[seu$seurat_clusters == stem_clust, , drop = FALSE])
vertices <- t(cds@principal_graph_aux[["UMAP"]]$dp_mst)
root_vertex <- rownames(vertices)[which.min(colSums((t(vertices) - centroid)^2))]
cds <- order_cells(cds, root_pr_nodes = root_vertex)
seu$pseudotime <- pseudotime(cds)

pdfout("[FigS6B] Pseudotime")
print(FeaturePlot(seu, features = "pseudotime") +
      scale_color_gradientn(colors = viridis::plasma(n = 10)))
dev.off()

saveRDS(seu, file.path(CDS, "cart1_cart2_scored.RDS"))
```