---
title: "Exp160 - Figure 4 and Figure S6"
author: "Koyama et al., Science Immunology (ady3001)"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: show
---

<!-- GENERATED from 01_fig4_figS6.R by Zenodo/manual_build/r_to_rmd.py.
     Every line of R below is byte-identical to that script, which is the
     version verified to reproduce the published objects cell-for-cell.
     Re-generate rather than hand-editing, or re-verify if you do edit. -->

Exp160 - Figure 4 and Figure S6
  cartridge1 (naive WT B6)  +  cartridge2 (24 h post-TBI WT B6)

Source: Exp160/rmd/Exp160_Final.Rmd, chunks
  "- loads RSEC data", "- analysis of clusters",
  "- expression markers MHCI and II", "- GSEA", "- custom gene sets",
  "pseudotime", "Lgr4 and Fgfbp1 - New ISC model"

Deviations from that Rmd, all forced, all listed in the methods page:
  * DefaultAssay(seu_sub) <- "RNA" is called AFTER seu_sub exists. The Rmd
    calls it one line before the object is created, which errors on a clean
    run (the chunk only worked because seu_sub was left over in the session).
  * The Monocle3 root node is chosen programmatically as the stem-cell
    cluster, which is what the Fig S6B legend states was used ("using stem
    cell cluster as root node"). The Rmd calls order_cells(cds) with no
    arguments, which opens a Shiny window and cannot run unattended.
  * set.seed(1234) is called before each processing block, as the Rmd does,
    but Seurat's and monocle3's own per-function seed defaults are left alone -
    they are already fixed, and overriding them changes the answer (see
    load_bd.R).

Usage:
  Rscript 01_fig4_figS6.R [--source deposit|original|rebuilt] [--which RSEC|DBEC]

<!-- original shebang: `#!/usr/bin/env Rscript` -->

```{r setup}
suppressPackageStartupMessages({
  library(Seurat); library(ggplot2); library(dplyr); library(patchwork)
  library(scCustomize); library(ggpubr); library(Matrix)
})
```

## Setting up

**To run this file, edit one line — `DATA` in the chunk below.** Download the
processed matrices for this experiment from NCBI GEO
[GSE348009](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE348009) to
wherever you like, then point `DATA` at that folder. It must contain
the per-cartridge folders `cartridge1/` and `cartridge2/`.

Nothing else needs editing. Output is written to `ady3001_output/` beside this
file.

```{r paths}
## ===========================================================================
##  EDIT THIS ONE LINE - where you put the GEO download for Exp160
## ===========================================================================
DATA <- "~/GSE348009/Exp160"
## ===========================================================================
##  Nothing below needs editing.
## ---------------------------------------------------------------------------
## Find this file, so common/ resolves whatever the working directory is.
.here <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                  error = function(e) normalizePath(getwd()))
PKG  <- normalizePath(file.path(.here, ".."), mustWork = TRUE)   # the Koyama/ folder
DATA <- path.expand(DATA)
if (!dir.exists(DATA))
  stop("DATA does not exist: ", DATA,
       "\n  Edit the DATA line above to point at your GEO download.")

source(file.path(PKG, "common", "config.R"))

SOURCE <- "deposit"   # the GEO matrices. Downstream code reads this - leave it.
WHICH  <- "RSEC"      # the BD UMI correction the published analysis used

OUT <- file.path(file.path(.here, "ady3001_output"), "Exp160", paste0("out_", SOURCE, "_", WHICH))
FIG <- file.path(OUT, "figs"); CDS <- file.path(OUT, "cds")
## Create and VERIFY. dir.create(showWarnings = FALSE) hides a failure, and a
## failure here is silent until the first pdf() call dies hundreds of lines later.
for (d in c(FIG, CDS)) {
  dir.create(d, recursive = TRUE, showWarnings = FALSE)
  if (!dir.exists(d)) stop("could not create output directory: ", d)
}

theme_set(bd_theme())
rna_cols    <- bd_rna_cols()
clus_cols36 <- bd_clus_cols36
SEED <- 1234
pdfout <- function(name, w = 6, h = 6) pdf(file.path(FIG, paste0(name, ".pdf")), width = w, height = h)

cat("== Exp160 Fig 4 / S6 ==\n   source:", SOURCE, "\n   data  :", DATA, "\n")
```

## 1. load cart1 + cart2

```{r sec1-load-cart1-cart2}
dirs <- c(cart1 = file.path(DATA, if (SOURCE == "original") "cart1" else "cartridge1"),
          cart2 = file.path(DATA, if (SOURCE == "original") "cart2" else "cartridge2"))
loaded <- bd_load(dirs, which = WHICH, n_pabo = 4)
seu <- bd_seurat(loaded)
cat("   cells before filtering:", ncol(seu), "\n")
```

## 2. QC filter, then re-process

The Rmd wrote  percent.mt < 25  while commenting "This gives us 8,032 samples".
Those two statements are not compatible: nCount_RNA 1e3-5e4 with mt < 25 keeps
6,412 cells, not 8,032. The object the figure was actually made from settles it -
Exp160/cds/241108v2_processed.RDS holds exactly 8,032 cells and its percent.mt
ranges 0.1705 to 39.99, i.e. it was filtered at mt < 40, the same threshold the
Fig S7 and Fig S10 analyses use. The 25 was a stale edit and the Rmd was
corrected to 40 on 2026-10-06; re-running it then reproduced the published
object cell-for-cell (identical barcodes and clusters, 100.00% agreement).

```{r sec2-qc-filter-then-re-process}
MT_MAX <- 40
seu_sub <- subset(seu, subset = nCount_RNA > 1000 & nCount_RNA < 50000 & percent.mt < MT_MAX)
cat("   cells after filtering :", ncol(seu_sub),
    "   (241108v2_processed.RDS holds 8,032)\n")
if (ncol(seu_sub) != 8032)
  warning("cell count after QC is ", ncol(seu_sub), ", expected 8032")
seu_sub <- bd_process(seu_sub, dims = 1:40, resolution = 0.15, seed = SEED)
seu_sub <- bd_process_adt(seu_sub)
saveRDS(seu_sub, file.path(CDS, "cart1_cart2_processed.RDS"))
seu <- seu_sub

## QC panels (not published, kept as the run's own evidence)
pdfout("QC - violin by cartridge", 10, 4)
print(VlnPlot(seu, features = c("nFeature_RNA","log_RNA","logUMI","percent.mt"),
              group.by = "cart", ncol = 4, pt.size = 0))
dev.off()
```

## 3. Figure 4B

```{r sec3-figure-4b}
pdfout("[Fig4B] Initial Clustering, res=0.15, dim=40")
print(DimPlot(seu, cols = clus_cols36)); dev.off()
pdfout("[Fig4B] Initial Clustering by cart")
print(DimPlot(seu, group.by = "cart")); dev.off()

## cluster identity markers used to name the 7 clusters
pdfout("[Fig4B] Markers dotplot")
print(DotPlot(seu, features = c("Chgb","Cpe","Muc2","Lyz1","Mmp7","Trpm5",
                                "Pou2f3","Anpep","Epcam","Lgr5","Ccl25")) + RotatedAxis())
dev.off()
for (g in c("Chgb","Muc2","Trpm5","Lyz1","Mmp7","Pou2f3","Anpep","Epcam","Lgr5")) {
  pdfout(paste0("[Fig4B] marker ", g)); print(FeaturePlot_scCustom(seu, features = g)); dev.off()
}
```

## 4. Figure 4C - MSigDB C5 module scores

```{r sec4-figure-4c-msigdb-c5-module-scores}
suppressPackageStartupMessages({ library(msigdbr) })
m_t2g <- msigdbr(species = "Mus musculus", category = "C5") %>%
  dplyr::select(gs_name, gene_symbol)

my_comparisons <- list(c("0","1"),c("0","2"),c("0","3"),c("0","4"),c("0","5"),c("0","6"))

gs_panels <- list(
  list(set = "GOCC_MHC_CLASS_II_PROTEIN_COMPLEX", name = "MHC_CLASS_II_PROTEIN_COMPLEX", ymax = 5),
  list(set = "GOCC_MHC_CLASS_I_PROTEIN_COMPLEX",  name = "MHC_CLASS_I_PROTEIN_COMPLEX",  ymax = 2.5),
  list(set = "GOBP_ANTIGEN_PROCESSING_AND_PRESENTATION",
       name = "ANTIGEN_PROCESSING_AND_PRESENTATION", ymax = 1))

DefaultAssay(seu) <- "RNA"
for (p in gs_panels) {
  genes <- m_t2g[m_t2g$gs_name == p$set, ]$gene_symbol
  stopifnot(length(genes) > 0)
  set.seed(SEED)
  seu <- AddModuleScore(seu, features = list(genes), name = p$name)
  feat <- paste0(p$name, "1")
  pdfout(paste0("[Fig4C] ", p$set))
  print(FeaturePlot_scCustom(seu, features = feat, label = FALSE, pt.size = 1,
                             max.cutoff = "q99") + scale_color_gradientn(colors = rna_cols))
  dev.off()
  pdfout(paste0("[Fig4C] ", p$set, " violin"))
  print(VlnPlot(seu, features = feat, group.by = "seurat_clusters",
                col = clus_cols36, pt.size = 0) +
        stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
        stat_compare_means(comparisons = my_comparisons, method = "wilcox.test") +
        ylim(0, p$ymax))
  dev.off()
}
```

## 5. Figure 4D - surface protein

```{r sec5-figure-4d-surface-protein}
DefaultAssay(seu) <- "ADT"
for (ab in list(list(f = "I-A-I-E", ymax = 12), list(f = "H-2Kb", ymax = 14))) {
  pdfout(paste0("[Fig4D] ", ab$f))
  print(FeaturePlot_scCustom(seu, features = ab$f, max.cutoff = "q99")); dev.off()
  pdfout(paste0("[Fig4D] ", ab$f, " (violin)"))
  print(VlnPlot(seu, features = ab$f, group.by = "seurat_clusters",
                col = clus_cols36, pt.size = 0) +
        stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
        stat_compare_means(comparisons = my_comparisons, method = "wilcox.test") +
        ylim(0, ab$ymax))
  dev.off()
}
```

## 6. Figure S6A - published cell-subset signatures

Gene sets as listed in the Rmd, from:
  Grun 2015 PMID 26287467; Haber 2017 PMID 29144463; Biton 2018 PMID 30392957;
  Hadjsombati 2018 PMID 30021144; Wang 2019 PMID 31753849; Zhao 2020 PMID 31900405;
  Yin 2023 doi 10.1038/s41421-023-00578-4; Malagola 2024 (Atad2/Stmn1/Smarca5/Dek)

```{r sec6-figure-s6a-published-cell-subset-signatures}
custom_gene_sets <- list(
  Stem_Cell                = c("Lgr5","Olfm4","Slc12a2","Ascl2","Axin2","Gkn3"),
  Transit_Amplifying_Cells = c("Olfm4","Scl12a2"),
  Goblet                   = c("Muc2","Agr2","Clca3","Tff3","Ang4","Defa31"),
  Tuft                     = c("Dclk1","Trpm5","Gnat3","CD24a","Krt8","Krt18","Gfi1b",
                               "Il25","Ptgs1","Alox5","Ptprc","Chat","Rac2"),
  Paneth                   = c("Mptx1","Ang4","Defa31","Defa17","Defa24","Defa22","Lyz1","Itln1"),
  Enteroendocrine          = c("Reg4","Chga","Chgb","Tph1","Tac1","Neuro3"),
  Enterocytes              = c("Alpi","Apoa1","Apoa4","Fabp1"),
  Partial_IEC              = c("Top2a","Apol10a","Car4","Ube2c"),
  Malagola_2024            = c("Atad2","Stmn1","Smarca5","Dek"))

DefaultAssay(seu) <- "RNA"
for (nm in names(custom_gene_sets)) {
  set.seed(SEED)
  seu <- AddModuleScore(seu, features = list(custom_gene_sets[[nm]]), name = nm)
  pdfout(paste0("[FigS6A] Custom ", nm))
  print(FeaturePlot_scCustom(seu, features = paste0(nm, "1"), label = FALSE,
                             pt.size = 1, max.cutoff = "q99") +
        scale_color_gradientn(colors = rna_cols))
  dev.off()
}
```

## 7. Figure S6C - ISC / progenitor markers

```{r sec7-figure-s6c-isc-progenitor-markers}
for (g in c("Lgr5","Lgr4","Fgfbp1","H2-Aa","H2-Ab1")) {
  pdfout(paste0("[FigS6C] ", g)); print(FeaturePlot_scCustom(seu, features = g)); dev.off()
}
```

## 8. Figure S6D/E - dot plots

```{r sec8-figure-s6d-e-dot-plots}
DefaultAssay(seu) <- "RNA"
pdfout("[FigS6D] Markers GSEA dotplot", 4, 6)
print(DotPlot(seu, features = c("MHC_CLASS_II_PROTEIN_COMPLEX1",
                                "MHC_CLASS_I_PROTEIN_COMPLEX1",
                                "ANTIGEN_PROCESSING_AND_PRESENTATION1")) +
      scale_x_discrete(labels = function(x) substr(x, 1, 15)) + RotatedAxis())
dev.off()
DefaultAssay(seu) <- "ADT"
pdfout("[FigS6E] Markers ADT dotplot", 4, 6)
print(DotPlot(seu, features = c("I-A-I-E", "H-2Kb")) + RotatedAxis()); dev.off()
DefaultAssay(seu) <- "RNA"
```

## 9. Figure S6B - Monocle3 pseudotime

Root node: the legend says the stem-cell cluster is the root. The Rmd picks
it by hand in order_cells()'s Shiny window. Here the stem-cell cluster is the
one with the highest mean Stem_Cell module score, and the root is the graph
vertex closest to that cluster's UMAP centroid. Nothing else changes.

```{r sec9-figure-s6b-monocle3-pseudotime}
suppressPackageStartupMessages({ library(monocle3); library(SeuratWrappers) })
set.seed(SEED)
cds <- as.cell_data_set(seu)
cds <- cluster_cells(cds, resolution = 1e-3)   # monocle3's own default random_seed, as in the Rmd
cds <- learn_graph(cds, use_partition = TRUE, close_loop = TRUE)
cds@clusters@listData[["UMAP"]][["clusters"]] <- clusters(cds)

stem_score  <- tapply(seu$Stem_Cell1, seu$seurat_clusters, mean)
stem_clust  <- names(which.max(stem_score))
cat("   pseudotime root cluster (highest Stem_Cell score):", stem_clust,
    sprintf("(score %.3f)\n", max(stem_score)))
centroid <- colMeans(Embeddings(seu, "umap")[seu$seurat_clusters == stem_clust, , drop = FALSE])
vertices <- t(cds@principal_graph_aux[["UMAP"]]$dp_mst)
root_vertex <- rownames(vertices)[which.min(colSums((t(vertices) - centroid)^2))]
cds <- order_cells(cds, root_pr_nodes = root_vertex)
seu$pseudotime <- pseudotime(cds)

pdfout("[FigS6B] Pseudotime")
print(FeaturePlot(seu, features = "pseudotime") +
      scale_color_gradientn(colors = viridis::plasma(n = 10)))
dev.off()

saveRDS(seu, file.path(CDS, "cart1_cart2_scored.RDS"))
```

## 10. numbers the methods page quotes

```{r sec10-numbers-the-methods-page-quotes}
tab <- as.data.frame(table(cluster = seu$seurat_clusters, cartridge = seu$cart))
write.csv(tab, file.path(OUT, "cluster_by_cartridge.csv"), row.names = FALSE)
write.csv(data.frame(metric = c("cells_loaded","cells_after_filter","n_clusters"),
                     value  = c(ncol(loaded$rna), ncol(seu), nlevels(seu$seurat_clusters))),
          file.path(OUT, "counts.csv"), row.names = FALSE)
bd_session(file.path(OUT, "sessionInfo.txt"))
cat("== done ->", OUT, "\n")
```
