---
title: "Exp160 - Figure S7"
author: "Koyama et al., Science Immunology (ady3001)"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: show
---

<!-- GENERATED from 02_figS7.R by Zenodo/manual_build/r_to_rmd.py.
     Every line of R below is byte-identical to that script, which is the
     version verified to reproduce the published objects cell-for-cell.
     Re-generate rather than hand-editing, or re-verify if you do edit. -->

Exp160 - Figure S7
  cartridge2 (24 h post-TBI WT B6)  +  cartridge3 (24 h post-TBI MHC-II KO)

Source: Exp160/rmd/Exp160_Final.Rmd, chunks
  "add Cart 3 - re-embed with Cart 2 only",
  "add Cart 3 with cart 2 re-embed - custom gene set and other ADT/RNA analysis",
  "add Cart 3 with card 2 re-embed - GSEA",
  "add Cart 3 with card 2 re-embed - Propellar and DEGs"

QC threshold: percent.mt < 40, and resolution 0.1 here against 0.15
in the Fig 4 / S6 analysis.

NOTE: the Rmd read "< 25" for the Fig 4 / S6 analysis until it was
corrected on 2026-10-06. That value was STALE - the published Fig 4
object holds 8,032 cells with a percent.mt maximum of 39.994, and 25
yields 6,412. BOTH analyses used 40, and the Rmd now says so. See
01_original_code/KNOWN_ISSUES.md. The resolutions (0.1 / 0.15)
genuinely do differ and are correct as stated.

One forced deviation, listed in the methods page:
  * the volcano in S7E. The Rmd builds it as
      volcano_0_2 <- data.frame(cbind(rownames(test0_2),
                                      test0_2$p_val_adj < 0.05,
                                      test0_2$avg_log2FC,
                                      test0_1$p_val))      # <- test0_1
    test0_1 is never assigned anywhere in the Exp160 Rmd corpus, so the line
    raises "object 'test0_1' not found" on a clean run. Every other field of
    that data.frame comes from test0_2 (cluster 0 vs cluster 2), so the
    p-value column is taken from test0_2$p_val here.

Usage: Rscript 02_figS7.R [--source deposit|original|rebuilt] [--which RSEC|DBEC]

> **Knitting this file.** The code locates the package root by walking up
> for a `.ady3001_root` marker, so knit with `knit_root_dir` set to the
> package root, or run from a directory beneath it. Point `$ADY3001_DATA`
> at an unpacked GEO download; `tools/unpack_geo_download.sh` builds that
> tree. Results are written to `$ADY3001_OUT`, default `./ady3001_output`.

<!-- original shebang: `#!/usr/bin/env Rscript` -->

```{r setup}
suppressPackageStartupMessages({
  library(Seurat); library(ggplot2); library(dplyr); library(patchwork)
  library(scCustomize); library(ggpubr); library(Matrix); library(ggrepel)
})
```

## path resolution (Zenodo package)

These lines replace the hardcoded lab paths of the working copy. Every other
change to this file is also path-only; no analysis line differs from the
script that produced the published figures. The complete diff against the
working copy is shipped at 05_verification/path_rewrite.diff - read it rather
than taking this comment's word for it.

```{r path-resolution}
local({
  a <- commandArgs(FALSE)
  f <- sub("^--file=", "", a[grep("^--file=", a)])
  d <- if (length(f)) dirname(normalizePath(f[1])) else normalizePath(getwd())
  repeat {
    if (file.exists(file.path(d, ".ady3001_root"))) break
    p <- dirname(d)
    if (p == d) stop("cannot locate the package root (.ady3001_root marker)")
    d <- p
  }
  assign("PKG", d, envir = .GlobalEnv)
})
source(file.path(PKG, "02_figure_scripts", "common", "config.R"))

args <- commandArgs(trailingOnly = TRUE)
getopt <- function(f, d) { i <- match(f, args); if (is.na(i)) d else args[i + 1L] }
SOURCE <- getopt("--source", "deposit"); WHICH <- getopt("--which", "RSEC")

DATA <- switch(SOURCE,
  original = file.path(ROOT, "Exp160", "data"),
  rebuilt  = file.path(REB,  "Exp160", "01_bd_pipeline"),
  ## "deposit" = the CLEAN ROOM: nothing but the files actually on GEO.
  ## Set $ADY3001_DATA to point at your unpacked GEO download.
  deposit  = file.path(Sys.getenv("ADY3001_DATA", file.path(PKG, "data")), "Exp160", "data"),
  stop("--source must be original, rebuilt or deposit"))
ady_check_data(DATA, SOURCE)

OUT <- file.path(OUT_ROOT, "Exp160", paste0("outS7_", SOURCE, "_", WHICH))
FIG <- file.path(OUT, "figs"); CDS <- file.path(OUT, "cds")
## Create and VERIFY. dir.create(showWarnings = FALSE) hides a failure, and a
## failure here is silent until the first pdf() call dies hundreds of lines later.
for (d in c(FIG, CDS)) {
  dir.create(d, recursive = TRUE, showWarnings = FALSE)
  if (!dir.exists(d)) stop("could not create output directory: ", d)
}

theme_set(bd_theme()); rna_cols <- bd_rna_cols(); clus_cols36 <- bd_clus_cols36
SEED <- 1234
pdfout <- function(n, w = 6, h = 6) pdf(file.path(FIG, paste0(n, ".pdf")), width = w, height = h)

cat("== Exp160 Fig S7 ==\n   source:", SOURCE, "\n")

dirs <- c(cart2 = file.path(DATA, if (SOURCE == "original") "cart2" else "cartridge2"),
          cart3 = file.path(DATA, if (SOURCE == "original") "cart3" else "cartridge3"))
loaded <- bd_load(dirs, which = WHICH, n_pabo = 4)
seu <- bd_seurat(loaded)
cat("   cells before filtering:", ncol(seu), "\n")

seu <- subset(seu, subset = nCount_RNA > 1000 & nCount_RNA < 50000 & percent.mt < 40)
cat("   cells after filtering :", ncol(seu), "   (Rmd recorded 7,478 from 23,841)\n")
seu <- bd_process(seu, dims = 1:40, resolution = 0.1, seed = SEED)
seu <- bd_process_adt(seu)
saveRDS(seu, file.path(CDS, "cart2_cart3_processed.RDS"))
```

## S7B: embedding by group, by cluster, and cluster proportions

```{r s7b-embedding-by-group-by-cluster-and-cluster-proportions}
pdfout("[FigS7B] Cart2 v Cart3 by cart")
print(DimPlot(seu, group.by = "cart")); dev.off()
pdfout("[FigS7B] Cart2 v Cart3 clusters, res=0.1, dim=40")
print(DimPlot(seu, cols = clus_cols36)); dev.off()

## proportions bar - propeller / speckle, Phipson 2022 PMID 36005887
suppressPackageStartupMessages({ library(speckle); library(limma) })
pdf(file.path(FIG, "[FigS7B] Propeller_plots.pdf"), width = 3, height = 8)
print(plotCellTypeProps(clusters = seu$seurat_clusters, sample = seu$cart) +
      scale_fill_manual(values = clus_cols36[seq_len(nlevels(seu$seurat_clusters))]) +
      theme_classic())
dev.off()
```

## S7C: published cell-subset signatures

```{r s7c-published-cell-subset-signatures}
custom_gene_sets <- list(
  Stem_Cell                = c("Lgr5","Olfm4","Slc12a2","Ascl2","Axin2","Gkn3"),
  Transit_Amplifying_Cells = c("Olfm4","Scl12a2"),
  Goblet                   = c("Muc2","Agr2","Clca3","Tff3","Ang4","Defa31"),
  Tuft                     = c("Dclk1","Trpm5","Gnat3","CD24a","Krt8","Krt18","Gfi1b",
                               "Il25","Ptgs1","Alox5","Ptprc","Chat","Rac2"),
  Paneth                   = c("Mptx1","Ang4","Defa31","Defa17","Defa24","Defa22","Lyz1","Itln1"),
  Enteroendocrine          = c("Reg4","Chga","Chgb","Tph1","Tac1","Neuro3"),
  Enterocytes              = c("Alpi","Apoa1","Apoa4","Fabp1"),
  Partial_IEC              = c("Top2a","Apol10a","Car4","Ube2c"))

DefaultAssay(seu) <- "RNA"
for (nm in names(custom_gene_sets)) {
  set.seed(SEED)
  seu <- AddModuleScore(seu, features = list(custom_gene_sets[[nm]]), name = nm)
  pdfout(paste0("[FigS7C] Custom ", nm))
  print(FeaturePlot_scCustom(seu, features = paste0(nm, "1"), label = FALSE,
                             pt.size = 1, max.cutoff = "q99") +
        scale_color_gradientn(colors = rna_cols))
  dev.off()
}
```

## S7D: protein and gene-set panels, with per-cluster Wilcoxon

```{r s7d-protein-and-gene-set-panels-with-per-cluster-wilcoxon}
suppressPackageStartupMessages({ library(msigdbr) })
m_t2g <- msigdbr(species = "Mus musculus", category = "C5") %>%
  dplyr::select(gs_name, gene_symbol)

lv <- levels(seu$seurat_clusters)
my_comparisons <- lapply(setdiff(lv, "0"), function(x) c("0", x))
cat("   clusters:", paste(lv, collapse = ","), "\n")

gs_panels <- list(
  list(set = "GOCC_MHC_CLASS_II_PROTEIN_COMPLEX", name = "MHC_CLASS_II_PROTEIN_COMPLEX", ymax = 7),
  list(set = "GOCC_MHC_CLASS_I_PROTEIN_COMPLEX",  name = "MHC_CLASS_I_PROTEIN_COMPLEX",  ymax = 3),
  list(set = "GOBP_ANTIGEN_PROCESSING_AND_PRESENTATION",
       name = "ANTIGEN_PROCESSING_AND_PRESENTATION", ymax = 1.2))

for (p in gs_panels) {
  genes <- m_t2g[m_t2g$gs_name == p$set, ]$gene_symbol
  set.seed(SEED)
  seu <- AddModuleScore(seu, features = list(genes), name = p$name)
  feat <- paste0(p$name, "1")
  pdfout(paste0("[FigS7D] ", p$set))
  print(FeaturePlot_scCustom(seu, features = feat, label = FALSE, pt.size = 1,
                             max.cutoff = "q99") + scale_color_gradientn(colors = rna_cols))
  dev.off()
  pdfout(paste0("[FigS7D] ", p$set, " violin"))
  print(VlnPlot(seu, features = feat, group.by = "seurat_clusters",
                col = clus_cols36, pt.size = 0) +
        stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
        stat_compare_means(comparisons = my_comparisons, method = "wilcox.test") +
        ylim(0, p$ymax))
  dev.off()
}

DefaultAssay(seu) <- "ADT"
for (ab in c("I-A-I-E", "H-2Kb")) {
  pdfout(paste0("[FigS7D] ", ab))
  print(FeaturePlot_scCustom(seu, features = ab, max.cutoff = "q99")); dev.off()
  pdfout(paste0("[FigS7D] ", ab, " (violin)"))
  print(VlnPlot(seu, features = ab, group.by = "seurat_clusters",
                col = clus_cols36, pt.size = 0) +
        stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
        stat_compare_means(comparisons = my_comparisons, method = "wilcox.test"))
  dev.off()
}
pdfout("[FigS7D] Markers ADT dotplot", 4, 6)
print(DotPlot(seu, features = c("I-A-I-E", "H-2Kb")) + RotatedAxis()); dev.off()
DefaultAssay(seu) <- "RNA"
pdfout("[FigS7D] Markers GSEA dotplot", 4, 6)
print(DotPlot(seu, features = c("MHC_CLASS_II_PROTEIN_COMPLEX1",
                                "MHC_CLASS_I_PROTEIN_COMPLEX1",
                                "ANTIGEN_PROCESSING_AND_PRESENTATION1")) +
      scale_x_discrete(labels = function(x) substr(x, 1, 15)) + RotatedAxis())
dev.off()
```

## S7E: DEGs, cluster 0 (enterocyte WT) vs cluster 2 (enterocyte KO)

```{r s7e-degs-cluster-0-vs-cluster-2}
DefaultAssay(seu) <- "RNA"
stopifnot(all(c("0","2") %in% lv))
test0_2 <- FindMarkers(seu, ident.1 = "0", ident.2 = "2")
write.csv(test0_2, file.path(OUT, "DEGs_cluster0_vs_cluster2.csv"))

de <- data.frame(delabel        = rownames(test0_2),
                 diffexpressed  = test0_2$p_val_adj < 0.05,
                 log2FoldChange = test0_2$avg_log2FC,
                 pvalue         = test0_2$p_val)      # see header note re test0_1
de$pvalue[de$pvalue == 0] <- .Machine$double.xmin
neg_col <- "#00BFC4"; pos_col <- "#F8766D"
de_df <- de %>% filter(diffexpressed) %>%
  mutate(color = ifelse(log2FoldChange < 0, neg_col,
                 ifelse(log2FoldChange > 0, pos_col, "grey70")))
cat("   DE genes at padj<0.05:", nrow(de_df), "\n")

for (n in c(30, 150)) {
  top <- de_df %>% arrange(pvalue) %>% slice_head(n = n)
  pdfout(paste0("[FigS7E] DEG cluster 0v2 top ", n))
  print(ggplot(de_df, aes(x = log2FoldChange, y = -log10(pvalue), color = color)) +
          geom_point(size = 1.8) +
          ggrepel::geom_text_repel(data = top, aes(label = delabel), max.overlaps = Inf,
                                   box.padding = 0.4, point.padding = 0.2,
                                   segment.alpha = 0.5, seed = 123) +
          scale_color_identity(guide = "none") + theme_bw() +
          labs(x = "log2 fold-change", y = "-log10(p-value)"))
  dev.off()
}

pdf(file.path(FIG, "[FigS7E] top DEGs featureplots.pdf"), width = 12, height = 16)
print(FeaturePlot_scCustom(seu, features = c("Dmbt1","Ubd","Saa1","Reg3b","Reg3g","Rnf186"),
                           max.cutoff = "q99", colors_use = rna_cols, pt.size = 0.2))
dev.off()

saveRDS(seu, file.path(CDS, "cart2_cart3_scored.RDS"))
write.csv(as.data.frame(table(cluster = seu$seurat_clusters, cartridge = seu$cart)),
          file.path(OUT, "cluster_by_cartridge.csv"), row.names = FALSE)
bd_session(file.path(OUT, "sessionInfo.txt"))
cat("== done ->", OUT, "\n")
```
