---
title: "Exp160 - Figure S7"
author: "Koyama et al., Science Immunology"
output: html_document
---

# Exp160 — Figure S7

# Ileal intestinal epithelial cells from **cartridge2** (24 h post-TBI WT B6) and **cartridge3** (24 h post-TBI MHC-II KO), re-clustered together.

```{r setup}
suppressPackageStartupMessages({
  library(Seurat); library(ggplot2); library(dplyr); library(patchwork)
  library(scCustomize); library(ggpubr); library(Matrix); library(ggrepel)
})
```

## Setting up

**Edit one line — `DATA`.** Download this experiment's processed matrices
from NCBI GEO [GSE348009](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE348009)
to wherever you like and point `DATA` at that folder. It must contain
`cartridge2/` and `cartridge3/`.

Output is written to
`Exp160_figS7/` in the working directory.

```{r paths}
## ==========  EDIT THIS ONE LINE  ==========
DATA <- "~/GSE348009/Exp160"
## ==========================================

OUT <- "Exp160_figS7"
FIG <- file.path(OUT, "figs"); CDS <- file.path(OUT, "cds")
for (d in c(FIG, CDS)) dir.create(d, recursive = TRUE, showWarnings = FALSE)

theme_set(bd_theme()); rna_cols <- bd_rna_cols(); clus_cols36 <- bd_clus_cols36
SEED <- 1234
pdfout <- function(n, w = 6, h = 6) pdf(file.path(FIG, paste0(n, ".pdf")), width = w, height = h)

cat("== Exp160 Fig S7 ==\n")
```


## Helper functions
```{r functions}
suppressPackageStartupMessages({
  library(Matrix)
  library(data.table)
  library(Seurat)
})

## Path to the filtered per-cell molecule matrix for one cartridge.
## `which` is "RSEC" or "DBEC".
bd_matrix_path <- function(dir, which = c("RSEC", "DBEC")) {
  f <- list.files(dir, full.names = TRUE)
  hit <- grep(paste0("_", which, "_MolsPerCell\\.csv$"), f, value = TRUE)
  hit
}

## Number of AbSeq targets, counted from the file's own header line.
bd_n_pabo <- function(path) {
  con <- file(path, "r"); on.exit(close(con))
  repeat {
    line <- readLines(con, n = 1L)
    if (!startsWith(line, "#")) break
  }
  cols <- strsplit(line, ",", fixed = TRUE)[[1]]
  sum(endsWith(cols, "|pAbO"))
}

## Read one or more cartridges into a single sparse matrix.
##   dirs      named character vector: name = cart prefix, value = directory
##   n_pabo    expected number of AbSeq rows 
bd_load <- function(dirs, which = "RSEC", n_pabo = NULL) {
  paths <- vapply(dirs, bd_matrix_path, character(1), which = which)
  k <- unique(vapply(paths, bd_n_pabo, integer(1)))
  message(sprintf("  %s: %d cartridge(s), %d AbSeq targets", which, length(paths), k))

  dat <- lapply(paths, function(p) {
    message("  reading ", basename(p))
    t(as.matrix(data.table::fread(p), rownames = 1))
  })

  rn <- Reduce(intersect, lapply(dat, rownames))
  rn <- vapply(strsplit(rn, "\\|"), `[[`, character(1), 1L)
  dat <- lapply(dat, function(m) {
    rownames(m) <- vapply(strsplit(rownames(m), "\\|"), `[[`, character(1), 1L)
    m[rn, , drop = FALSE]
  })
  for (i in seq_along(dat))
    colnames(dat[[i]]) <- paste0(names(dirs)[i], "_", colnames(dat[[i]]))

  mat <- Matrix(do.call(cbind, dat), sparse = TRUE)
  colnames(mat) <- make.unique(colnames(mat))
  list(rna = mat[(k + 1L):nrow(mat), , drop = FALSE],
       adt = mat[1L:k, , drop = FALSE],
       n_pabo = k)
}

## Build the Seurat object: RNA counts, the ADT (protein) assay, a cartridge
## label per cell, mitochondrial percentage and log10 UMI counts.
bd_seurat <- function(loaded) {
  seu <- CreateSeuratObject(counts = loaded$rna)
  seu[["ADT"]] <- CreateAssayObject(
    counts = loaded$adt[, match(colnames(loaded$adt), Cells(seu)), drop = FALSE])
  seu$cart <- vapply(strsplit(Cells(seu), "_"), `[[`, character(1), 1L)
  seu[["percent.mt"]] <- PercentageFeatureSet(seu, pattern = "^mt-")
  seu$logUMI  <- log10(seu$nCount_RNA)
  seu$log_RNA <- log10(seu$nCount_RNA)
  seu
}

## LogNormalize -> 3000 variable features -> scale -> PCA(50) -> neighbours,
## clusters, UMAP.
bd_process <- function(seu, dims, resolution, seed = 1234) {
  set.seed(seed)
  DefaultAssay(seu) <- "RNA"
  seu <- NormalizeData(seu, normalization.method = "LogNormalize", scale.factor = 10000)
  seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 3000)
  seu <- ScaleData(seu)
  seu <- RunPCA(seu, features = VariableFeatures(object = seu), npcs = 50)
  seu <- FindNeighbors(seu, dims = dims)
  seu <- FindClusters(seu, resolution = resolution)
  seu <- RunUMAP(seu, dims = dims)
  seu
}

## CLR normalisation of the protein assay.
bd_process_adt <- function(seu) {
  DefaultAssay(seu) <- "ADT"
  seu <- NormalizeData(seu, normalization.method = "CLR")
  seu <- ScaleData(seu)
  DefaultAssay(seu) <- "RNA"
  seu
}

## The continuous expression palette: ArchR::paletteContinuous(n=8)[c(1:3,6:8)].
bd_rna_cols <- function() {
  ArchR::paletteContinuous(n = 8)[c(1:3, 6:8)]
}

bd_clus_cols36 <- c(
  "#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b",
  "#e377c2", "#7f7f7f", "#bcbd22", "#17becf", "#aec7e8", "#ffbb78",
  "#98df8a", "#ff9896", "#c5b0d5", "#c49c94", "#f7b6d2", "#c7c7c7",
  "#dbdb8d", "#9edae5", "#ad494a", "#8c6d31", "#e7ba52", "#17becf",
  "#1f78b4", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#cab2d6",
  "#6a3d9a", "#ff8c00", "#b15928", "#41ab5d", "#f03b20", "#807dba")

bd_theme <- function() {
  library(ggplot2)
  theme_bw(base_size = 14) +
    theme(panel.background   = element_rect(fill = "transparent", colour = NA),
          panel.grid.minor   = element_blank(),
          panel.grid.major   = element_blank(),
          legend.background  = element_rect(fill = "transparent"),
          legend.box.background = element_rect(fill = "transparent"),
          legend.key         = element_rect(fill = "transparent", colour = NA),
          plot.background    = element_rect(fill = "transparent", colour = NA))
}

## Write the R and package versions beside the results.
bd_session <- function(path) {
  writeLines(c(capture.output(sessionInfo()), "",
               paste("run at", format(Sys.time(), "%Y-%m-%d %H:%M:%S"))), path)
}
```

## 1. load cart2 + cart3

```{r sec1-load-cart2-cart3}
dirs <- c(cart2 = file.path(DATA, "cartridge2"),
          cart3 = file.path(DATA, "cartridge3"))
loaded <- bd_load(dirs, which = "RSEC", n_pabo = 4)
seu <- bd_seurat(loaded)
cat("   cells before filtering:", ncol(seu), "\n")
```

## 2. QC filter, then re-process

```{r sec2-qc-filter-then-re-process}

seu <- bd_process(seu, dims = 1:40, resolution = 0.5, seed = SEED)
cat("   first pass:", ncol(seu), "cells,",
    nlevels(seu$seurat_clusters), "clusters at resolution 0.5\n")
seu <- subset(seu, subset = nCount_RNA > 1000 & nCount_RNA < 50000 & percent.mt < 40)
cat("   cells after filtering :", ncol(seu), "   (expected 7,478 from 23,841)\n")
seu <- bd_process(seu, dims = 1:40, resolution = 0.1, seed = SEED)
seu <- bd_process_adt(seu)
saveRDS(seu, file.path(CDS, "cart2_cart3_processed.RDS"))
```

## S7B: embedding by group, by cluster, and cluster proportions

```{r s7b-embedding-by-group-by-cluster-and-cluster-proportions}
pdfout("[FigS7B] Cart2 v Cart3 by cart")
print(DimPlot(seu, group.by = "cart")); dev.off()
pdfout("[FigS7B] Cart2 v Cart3 clusters, res=0.1, dim=40")
print(DimPlot(seu, cols = clus_cols36)); dev.off()

## proportions bar - propeller / speckle, Phipson 2022 PMID 36005887
suppressPackageStartupMessages({ library(speckle); library(limma) })
pdf(file.path(FIG, "[FigS7B] Propeller_plots.pdf"), width = 3, height = 8)
print(plotCellTypeProps(clusters = seu$seurat_clusters, sample = seu$cart) +
      scale_fill_manual(values = clus_cols36[seq_len(nlevels(seu$seurat_clusters))]) +
      theme_classic())
dev.off()
```

## S7C: published cell-subset signatures

```{r s7c-published-cell-subset-signatures}
custom_gene_sets <- list(
  Stem_Cell                = c("Lgr5","Olfm4","Slc12a2","Ascl2","Axin2","Gkn3"),
  Goblet                   = c("Muc2","Agr2","Clca3","Tff3","Ang4","Defa31"),
  Tuft                     = c("Dclk1","Trpm5","Gnat3","CD24a","Krt8","Krt18","Gfi1b",
                               "Il25","Ptgs1","Alox5","Ptprc","Chat","Rac2"),
  Paneth                   = c("Mptx1","Ang4","Defa31","Defa17","Defa24","Defa22","Lyz1","Itln1"),
  Enteroendocrine          = c("Reg4","Chga","Chgb","Tph1","Tac1","Neuro3"),
  Enterocytes              = c("Alpi","Apoa1","Apoa4","Fabp1"),
  Partial_IEC              = c("Top2a","Apol10a","Car4","Ube2c"))

DefaultAssay(seu) <- "RNA"
for (nm in names(custom_gene_sets)) {
  set.seed(SEED)
  seu <- AddModuleScore(seu, features = list(custom_gene_sets[[nm]]), name = nm)
  pdfout(paste0("[FigS7C] Custom ", nm))
  print(FeaturePlot_scCustom(seu, features = paste0(nm, "1"), label = FALSE,
                             pt.size = 1, max.cutoff = "q99") +
        scale_color_gradientn(colors = rna_cols))
  dev.off()
}
```

## S7D: protein and gene-set panels, with per-cluster Wilcoxon

```{r s7d-protein-and-gene-set-panels-with-per-cluster-wilcoxon}
options(warn = -1)
suppressPackageStartupMessages({ library(msigdbr) })
m_t2g <- msigdbr(species = "Mus musculus", category = "C5") %>%
  dplyr::select(gs_name, gene_symbol)

lv <- levels(seu$seurat_clusters)
my_comparisons <- lapply(setdiff(lv, "0"), function(x) c("0", x))
cat("   clusters:", paste(lv, collapse = ","), "\n")

gs_panels <- list(
  list(set = "GOCC_MHC_CLASS_II_PROTEIN_COMPLEX", name = "MHC_CLASS_II_PROTEIN_COMPLEX", ymax = 7),
  list(set = "GOCC_MHC_CLASS_I_PROTEIN_COMPLEX",  name = "MHC_CLASS_I_PROTEIN_COMPLEX",  ymax = 3),
  list(set = "GOBP_ANTIGEN_PROCESSING_AND_PRESENTATION",
       name = "ANTIGEN_PROCESSING_AND_PRESENTATION", ymax = 1.2))

for (p in gs_panels) {
  genes <- m_t2g[m_t2g$gs_name == p$set, ]$gene_symbol
  set.seed(SEED)
  seu <- AddModuleScore(seu, features = list(genes), name = p$name)
  feat <- paste0(p$name, "1")
  pdfout(paste0("[FigS7D] ", p$set))
  print(FeaturePlot_scCustom(seu, features = feat, label = FALSE, pt.size = 1,
                             max.cutoff = "q99") + scale_color_gradientn(colors = rna_cols))
  dev.off()
  pdfout(paste0("[FigS7D] ", p$set, " violin"))
  print(VlnPlot(seu, features = feat, group.by = "seurat_clusters",
                col = clus_cols36, pt.size = 0) +
        stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
        stat_compare_means(comparisons = my_comparisons, method = "wilcox.test") +
        ylim(0, p$ymax))
  dev.off()
}

DefaultAssay(seu) <- "ADT"
for (ab in c("I-A-I-E", "H-2Kb")) {
  pdfout(paste0("[FigS7D] ", ab))
  print(FeaturePlot_scCustom(seu, features = ab, max.cutoff = "q99")); dev.off()
  pdfout(paste0("[FigS7D] ", ab, " (violin)"))
  print(VlnPlot(seu, features = ab, group.by = "seurat_clusters",
                col = clus_cols36, pt.size = 0) +
        stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
        stat_compare_means(comparisons = my_comparisons, method = "wilcox.test"))
  dev.off()
}
```

## S7E: DEGs, cluster 0 (enterocyte WT) vs cluster 2 (enterocyte KO)

```{r s7e-degs-cluster-0-vs-cluster-2}
DefaultAssay(seu) <- "RNA"
stopifnot(all(c("0","2") %in% lv))
test0_2 <- FindMarkers(seu, ident.1 = "0", ident.2 = "2")
write.csv(test0_2, file.path(OUT, "DEGs_cluster0_vs_cluster2.csv"))

de <- data.frame(delabel        = rownames(test0_2),
                 diffexpressed  = test0_2$p_val_adj < 0.05,
                 log2FoldChange = test0_2$avg_log2FC,
                 pvalue         = test0_2$p_val)      # see header note re test0_1
de$pvalue[de$pvalue == 0] <- .Machine$double.xmin
neg_col <- "#00BFC4"; pos_col <- "#F8766D"
de_df <- de %>% filter(diffexpressed) %>%
  mutate(color = ifelse(log2FoldChange < 0, neg_col,
                 ifelse(log2FoldChange > 0, pos_col, "grey70")))
cat("   DE genes at padj<0.05:", nrow(de_df), "\n")

for (n in c(30, 150)) {
  top <- de_df %>% arrange(pvalue) %>% slice_head(n = n)
  pdfout(paste0("[FigS7E] DEG cluster 0v2 top ", n))
  print(ggplot(de_df, aes(x = log2FoldChange, y = -log10(pvalue), color = color)) +
          geom_point(size = 1.8) +
          ggrepel::geom_text_repel(data = top, aes(label = delabel), max.overlaps = Inf,
                                   box.padding = 0.4, point.padding = 0.2,
                                   segment.alpha = 0.5, seed = 123) +
          scale_color_identity(guide = "none") + theme_bw() +
          labs(x = "log2 fold-change", y = "-log10(p-value)"))
  dev.off()
}

pdf(file.path(FIG, "[FigS7E] top DEGs featureplots.pdf"), width = 12, height = 16)
print(FeaturePlot_scCustom(seu, features = c("Dmbt1","Ubd","Saa1","Reg3b","Reg3g","Rnf186"),
                           max.cutoff = "q99", colors_use = rna_cols, pt.size = 0.2))
dev.off()

saveRDS(seu, file.path(CDS, "cart2_cart3_scored.RDS"))
write.csv(as.data.frame(table(cluster = seu$seurat_clusters, cartridge = seu$cart)),
          file.path(OUT, "cluster_by_cartridge.csv"), row.names = FALSE)
bd_session(file.path(OUT, "sessionInfo.txt"))
cat("== done ->", OUT, "\n")
```
