---
title: "Exp649 - Figure S10 (QC, all cells)"
author: "Koyama et al., Science Immunology (ady3001)"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: show
---

<!-- GENERATED from 01_figS10_qc.R by Zenodo/manual_build/r_to_rmd.py.
     Every line of R below is byte-identical to that script, which is the
     version verified to reproduce the published objects cell-for-cell.
     Re-generate rather than hand-editing, or re-verify if you do edit. -->

Exp649 - Figure S10  (all sorted cells, before the ISC subset is taken)
  Cart1 + Cart2, day 7 after BMT, Marilyn T -> Ifngr1 dISC

Source: Exp649/rmd/Exp649_Final.Rmd, chunks
  "- loads RSEC data" and the first half of "- isolate ISCs"

The four experimental groups are NOT BD sample tags. Exp649's own metrics
header records "Sample Tag Version: None". The groups are carried by four
different CD326 (EpCAM, clone G8.8) AbSeq oligos, one per sorted population:

    CD326|Epcam|AMM2065|pAbO             catalog 940169   -> Female_IFNGRpos
    CD326:G8.8-AMM2281|Epcam|AMM2281     catalog 460165   -> Female_IFNGRneg
    CD326:G8.8-AMM2294|Epcam|AMM2294     catalog 460163   -> Male_IFNGRpos
    CD326:G8.8-AMM2295|Epcam|AMM2295     catalog 460164   -> Male_IFNGRneg

That oligo -> group mapping is the one place in this whole rebuild that is
not recoverable from any file on disk. It is transcribed from the Rmd, where
it is implicit in the order of the Ab1..Ab4 columns and the order of the
group names. It has to be confirmed against Motoko's staining record before
the GEO sample table is written. Nothing downstream can detect if it is
wrong - the clusters would simply carry the wrong labels.

Usage: Rscript 01_figS10_qc.R [--source deposit|original|rebuilt] [--which RSEC|DBEC]

> **Knitting this file.** The code locates the package root by walking up
> for a `.ady3001_root` marker, so knit with `knit_root_dir` set to the
> package root, or run from a directory beneath it. Point `$ADY3001_DATA`
> at an unpacked GEO download; `tools/unpack_geo_download.sh` builds that
> tree. Results are written to `$ADY3001_OUT`, default `./ady3001_output`.

<!-- original shebang: `#!/usr/bin/env Rscript` -->

```{r setup}
suppressPackageStartupMessages({
  library(Seurat); library(ggplot2); library(dplyr); library(scCustomize)
  library(ggpubr); library(Matrix)
})
```

## path resolution (Zenodo package)

These lines replace the hardcoded lab paths of the working copy. Every other
change to this file is also path-only; no analysis line differs from the
script that produced the published figures. The complete diff against the
working copy is shipped at 05_verification/path_rewrite.diff - read it rather
than taking this comment's word for it.

```{r path-resolution}
local({
  a <- commandArgs(FALSE)
  f <- sub("^--file=", "", a[grep("^--file=", a)])
  d <- if (length(f)) dirname(normalizePath(f[1])) else normalizePath(getwd())
  repeat {
    if (file.exists(file.path(d, ".ady3001_root"))) break
    p <- dirname(d)
    if (p == d) stop("cannot locate the package root (.ady3001_root marker)")
    d <- p
  }
  assign("PKG", d, envir = .GlobalEnv)
})
source(file.path(PKG, "02_figure_scripts", "common", "config.R"))

args <- commandArgs(trailingOnly = TRUE)
getopt <- function(f, d) { i <- match(f, args); if (is.na(i)) d else args[i + 1L] }
SOURCE <- getopt("--source", "deposit"); WHICH <- getopt("--which", "RSEC")

DATA <- switch(SOURCE,
  original = file.path(ROOT, "Exp649", "data"),
  rebuilt  = file.path(REB,  "Exp649", "01_bd_pipeline"),
  ## "deposit" = the CLEAN ROOM: nothing but the files actually on GEO.
  ## Set $ADY3001_DATA to point at your unpacked GEO download.
  deposit  = file.path(Sys.getenv("ADY3001_DATA", file.path(PKG, "data")), "Exp649", "data"),
  stop("--source must be original, rebuilt or deposit"))
ady_check_data(DATA, SOURCE)

OUT <- file.path(OUT_ROOT, "Exp649", paste0("outS10_", SOURCE, "_", WHICH))
FIG <- file.path(OUT, "figs"); CDS <- file.path(OUT, "cds")
## Create and VERIFY. dir.create(showWarnings = FALSE) hides a failure, and a
## failure here is silent until the first pdf() call dies hundreds of lines later.
for (d in c(FIG, CDS)) {
  dir.create(d, recursive = TRUE, showWarnings = FALSE)
  if (!dir.exists(d)) stop("could not create output directory: ", d)
}

theme_set(bd_theme()); rna_cols <- bd_rna_cols(); clus_cols36 <- bd_clus_cols36
SEED <- 1234
pdfout <- function(n, w = 6, h = 6) pdf(file.path(FIG, paste0(n, ".pdf")), width = w, height = h)

cat("== Exp649 Fig S10 (QC / all cells) ==\n   source:", SOURCE, "\n")

dirs <- c(cart1 = file.path(DATA, if (SOURCE == "original") "cart1" else "Cart1"),
          cart2 = file.path(DATA, if (SOURCE == "original") "cart2" else "Cart2"))
loaded <- bd_load(dirs, which = WHICH, n_pabo = 7)
seu <- bd_seurat(loaded)
cat("   cells before filtering:", ncol(seu), "\n")

## QC filter as in the Rmd: nCount 1e3-5e4, percent.mt < 40
seu <- subset(seu, subset = nCount_RNA > 1000 & nCount_RNA < 50000 & percent.mt < 40)
cat("   cells after filtering :", ncol(seu), "   (Rmd recorded 10,363)\n")
seu <- bd_process(seu, dims = 1:40, resolution = 0.5, seed = SEED)
seu <- bd_process_adt(seu)
```

## demultiplex the four CD326 oligos

```{r demultiplex-the-four-cd326-oligos}
ab_cols <- c(Ab1 = "CD326",
             Ab2 = "CD326:G8.8-AMM2281",
             Ab3 = "CD326:G8.8-AMM2294",
             Ab4 = "CD326:G8.8-AMM2295")
group_of <- c(Ab1 = "Female_IFNGRpos", Ab2 = "Female_IFNGRneg",
              Ab3 = "Male_IFNGRpos",   Ab4 = "Male_IFNGRneg")

adt <- GetAssayData(seu, assay = "ADT", layer = "data")
stopifnot(all(ab_cols %in% rownames(adt)))
pos <- sapply(ab_cols, function(a) adt[a, ] > 1)     # Rmd threshold: CLR value > 1
colnames(pos) <- names(ab_cols)

## density of each tag, the plot the Rmd draws before choosing the threshold
for (nm in names(ab_cols)) {
  d <- data.frame(value = adt[ab_cols[[nm]], ])
  pdfout(paste0("[QC] ADT density ", nm, " ", ab_cols[[nm]]), 5, 4)
  print(ggplot(d, aes(x = value)) + geom_density() +
        geom_vline(xintercept = 1, linetype = 2) +
        labs(title = paste0(nm, " = ", ab_cols[[nm]], " -> ", group_of[[nm]]),
             x = "CLR-normalised ADT"))
  dev.off()
}

n_tag <- rowSums(pos)
seu$multiplet    <- n_tag > 1
seu$Ab_not_found <- n_tag < 1
cat(sprintf("   multiplets: %d   no-tag: %d   singlets: %d\n",
            sum(seu$multiplet), sum(seu$Ab_not_found), sum(n_tag == 1)))
pdfout("[QC] multiplet"); print(DimPlot(seu, group.by = "multiplet")); dev.off()
pdfout("[QC] Ab_not_found"); print(DimPlot(seu, group.by = "Ab_not_found")); dev.off()

seu <- seu[, !seu$multiplet & !seu$Ab_not_found]
cat("   cells after de-multiplexing:", ncol(seu), "   (Rmd recorded 9,135 from 10,363)\n")
seu$group <- unname(group_of[colnames(pos)[apply(pos[Cells(seu), , drop = FALSE], 1, which.max)]])
print(table(seu$group, seu$cart))
pdfout("[QC] groups"); print(DimPlot(seu, group.by = "group")); dev.off()

saveRDS(seu, file.path(CDS, "all_cells_processed.RDS"))
```

## Figure S10

```{r figure-s10}
DefaultAssay(seu) <- "ADT"
pdfout("[FigS10A] QC Clusters")
print(DimPlot(seu, label = TRUE, label.size = 5, repel = TRUE, label.box = TRUE) +
      scale_fill_manual(values = rep("white", nlevels(seu$seurat_clusters)))); dev.off()

DefaultAssay(seu) <- "RNA"
for (g in c("Lgr5", "Muc2", "Trpm5")) {
  pdfout(paste0("[FigS10B-D] QC ", g)); print(FeaturePlot_scCustom(seu, features = g)); dev.off()
}

## S10B: Lgr5 across clusters, with the comparisons the legend names -
## cluster 8 against clusters 5, 6 and 9.
##
## Those numbers are used directly, but only after checking that this run's
## clusters really are the published ones. The check is the biology the Fig S10
## legend states: clusters 5, 6, 8 are the Muc2+ (goblet) clusters and 9 is the
## lowest-Lgr5 cluster. If both hold, the numbering is the published numbering and
## 8 vs 5/6/9 is the published comparison. If either fails, the run says so loudly
## and falls back to the measured rule rather than plotting the wrong clusters.
##
## (An earlier version read "cluster 8 (fourth lowest expression cluster)" in the
## legend as the selection RULE and picked the fourth-lowest Lgr5 cluster. That is
## a description of cluster 8, not a rule: it selected cluster 10 - the Trpm5+ tuft
## cluster - because 8 and 10 have nearly equal mean Lgr5.)
mean_lgr5 <- tapply(FetchData(seu, "Lgr5")[, 1], seu$seurat_clusters, mean)
mean_muc2 <- tapply(FetchData(seu, "Muc2")[, 1], seu$seurat_clusters, mean)
ord <- names(sort(mean_lgr5))
gapcut <- function(v) { sv <- sort(v); g <- diff(sv); if (!length(g)) return(Inf)
                        sv[which.max(g)] + max(g) / 2 }
goblet  <- names(mean_muc2)[mean_muc2 > gapcut(mean_muc2)]
lowest  <- ord[1]
cat("   Lgr5 rank (low -> high):", paste(ord, collapse = " "), "\n")
cat("   Muc2-high (goblet)     :", paste(sort(goblet), collapse = ","),
    "   (published: 5,6,8)\n")
cat("   lowest-Lgr5 cluster    :", lowest, "   (published: 9)\n")

as_published <- setequal(goblet, c("5","6","8")) && identical(lowest, "9")
if (as_published) {
  cat("   cluster numbering matches the published object; using 8 vs 5/6/9\n")
  ref <- "8"; others <- c("5","6","9")
} else {
  warning("cluster numbering does NOT match the published object - ",
          "S10B comparisons are derived, not the published ones")
  cat("   *** numbering differs; falling back to measured rule ***\n")
  ref <- ord[4]; others <- ord[1:3]
}
cat("   S10B comparisons: ", ref, " vs ", paste(others, collapse = "/"), "\n", sep = "")
my_comparisons <- lapply(others, function(x) c(x, ref))

pdfout("[FigS10B] QC Lgr5 Exp Level")
print(VlnPlot(seu, features = "Lgr5", pt.size = 0) +
      stat_summary(geom = "point", size = 10, colour = "black", shape = 95) +
      stat_compare_means(comparisons = my_comparisons, method = "wilcox.test") +
      ylim(0, 7)); dev.off()

pdfout("[FigS10C] mean Lgr5 Muc2 Trpm5", 7, 5)
mm <- FetchData(seu, c("Lgr5","Muc2","Trpm5"))
mm$cluster <- seu$seurat_clusters
mm <- mm %>% tidyr::pivot_longer(-cluster, names_to = "gene", values_to = "expr") %>%
  group_by(cluster, gene) %>% summarise(mean = mean(expr), .groups = "drop")
print(ggplot(mm, aes(cluster, mean, fill = gene)) +
      geom_col(position = "dodge") + theme_classic() + labs(y = "mean expression"))
dev.off()
write.csv(mm, file.path(OUT, "mean_Lgr5_Muc2_Trpm5_by_cluster.csv"), row.names = FALSE)

pdfout("[FigS10D] DotPlot Lgr5 Muc2 Trpm5", 6, 6)
print(DotPlot(seu, features = c("Lgr5","Muc2","Trpm5")) + RotatedAxis()); dev.off()

write.csv(data.frame(cluster = names(mean_lgr5), mean_Lgr5 = as.numeric(mean_lgr5)),
          file.path(OUT, "mean_Lgr5_by_cluster.csv"), row.names = FALSE)
write.csv(as.data.frame(table(cluster = seu$seurat_clusters, group = seu$group)),
          file.path(OUT, "cluster_by_group.csv"), row.names = FALSE)
bd_session(file.path(OUT, "sessionInfo.txt"))
cat("== done ->", OUT, "\n")
```
