---
title: "Exp649 - Figure S10 (QC, all cells)"
author: "Koyama et al., Science Immunology"
output: html_document
---

# Exp649 — Figure S10

# Ileal intestinal epithelial cells from **Cart1** and **Cart2**, day 7 after BMT, Marilyn T cells into Ifngr1 ΔISC recipients. All sorted cells, before the ISC subset is taken.

The four experimental groups arefour different CD326 (EpCAM, clone G8.8) AbSeq oligos, one per sorted
population:

    CD326|Epcam|AMM2065|pAbO             catalog 940169    Female IFNgR-positive
    CD326:G8.8-AMM2281|Epcam|AMM2281     catalog 460165    Female IFNgR-negative
    CD326:G8.8-AMM2294|Epcam|AMM2294     catalog 460163    Male   IFNgR-positive
    CD326:G8.8-AMM2295|Epcam|AMM2295     catalog 460164    Male   IFNgR-negative

```{r setup}
suppressPackageStartupMessages({
  library(Seurat); library(ggplot2); library(dplyr); library(scCustomize)
  library(ggpubr); library(Matrix)
})
```

## Setting up

**Edit one line — `DATA`.** Download this experiment's processed matrices
from NCBI GEO [GSE348009](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE348009)
to wherever you like and point `DATA` at that folder. It must contain
`Cart1/` and `Cart2/`.

Output is written to
`Exp649_figS10/` in the working directory.

```{r paths}
## ==========  EDIT THIS ONE LINE  ==========
DATA <- "~/GSE348009/Exp649"
## ==========================================

OUT <- "Exp649_figS10"
FIG <- file.path(OUT, "figs"); CDS <- file.path(OUT, "cds")
for (d in c(FIG, CDS)) dir.create(d, recursive = TRUE, showWarnings = FALSE)

theme_set(bd_theme()); rna_cols <- bd_rna_cols(); clus_cols36 <- bd_clus_cols36
SEED <- 1234
pdfout <- function(n, w = 6, h = 6) pdf(file.path(FIG, paste0(n, ".pdf")), width = w, height = h)

cat("== Exp649 Fig S10 (QC / all cells) ==\n")
```


## Helper functions
```{r functions}
suppressPackageStartupMessages({
  library(Matrix)
  library(data.table)
  library(Seurat)
})

## Path to the filtered per-cell molecule matrix for one cartridge.
## `which` is "RSEC" or "DBEC".
bd_matrix_path <- function(dir, which = c("RSEC", "DBEC")) {
  f <- list.files(dir, full.names = TRUE)
  hit <- grep(paste0("_", which, "_MolsPerCell\\.csv$"), f, value = TRUE)
  hit
}

## Number of AbSeq targets, counted from the file's own header line.
bd_n_pabo <- function(path) {
  con <- file(path, "r"); on.exit(close(con))
  repeat {
    line <- readLines(con, n = 1L)
    if (!startsWith(line, "#")) break
  }
  cols <- strsplit(line, ",", fixed = TRUE)[[1]]
  sum(endsWith(cols, "|pAbO"))
}

## Read one or more cartridges into a single sparse matrix.
##   dirs      named character vector: name = cart prefix, value = directory
##   n_pabo    expected number of AbSeq rows
bd_load <- function(dirs, which = "RSEC", n_pabo = NULL) {
  paths <- vapply(dirs, bd_matrix_path, character(1), which = which)
  k <- unique(vapply(paths, bd_n_pabo, integer(1)))
  message(sprintf("  %s: %d cartridge(s), %d AbSeq targets", which, length(paths), k))

  dat <- lapply(paths, function(p) {
    message("  reading ", basename(p))
    t(as.matrix(data.table::fread(p), rownames = 1))
  })

  rn <- Reduce(intersect, lapply(dat, rownames))
  rn <- vapply(strsplit(rn, "\\|"), `[[`, character(1), 1L)
  dat <- lapply(dat, function(m) {
    rownames(m) <- vapply(strsplit(rownames(m), "\\|"), `[[`, character(1), 1L)
    m[rn, , drop = FALSE]
  })
  for (i in seq_along(dat))
    colnames(dat[[i]]) <- paste0(names(dirs)[i], "_", colnames(dat[[i]]))

  mat <- Matrix(do.call(cbind, dat), sparse = TRUE)
  colnames(mat) <- make.unique(colnames(mat))
  list(rna = mat[(k + 1L):nrow(mat), , drop = FALSE],
       adt = mat[1L:k, , drop = FALSE],
       n_pabo = k)
}

## Build the Seurat object: RNA counts, the ADT (protein) assay, a cartridge
## label per cell, mitochondrial percentage and log10 UMI counts.
bd_seurat <- function(loaded) {
  seu <- CreateSeuratObject(counts = loaded$rna)
  seu[["ADT"]] <- CreateAssayObject(
    counts = loaded$adt[, match(colnames(loaded$adt), Cells(seu)), drop = FALSE])
  seu$cart <- vapply(strsplit(Cells(seu), "_"), `[[`, character(1), 1L)
  seu[["percent.mt"]] <- PercentageFeatureSet(seu, pattern = "^mt-")
  seu$logUMI  <- log10(seu$nCount_RNA)
  seu$log_RNA <- log10(seu$nCount_RNA)
  seu
}

## LogNormalize -> 3000 variable features -> scale -> PCA(50) -> neighbours,
## clusters, UMAP.
bd_process <- function(seu, dims, resolution, seed = 1234) {
  set.seed(seed)
  DefaultAssay(seu) <- "RNA"
  seu <- NormalizeData(seu, normalization.method = "LogNormalize", scale.factor = 10000)
  seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 3000)
  seu <- ScaleData(seu)
  seu <- RunPCA(seu, features = VariableFeatures(object = seu), npcs = 50)
  seu <- FindNeighbors(seu, dims = dims)
  seu <- FindClusters(seu, resolution = resolution)
  seu <- RunUMAP(seu, dims = dims)
  seu
}

## CLR normalisation of the protein assay.
bd_process_adt <- function(seu) {
  DefaultAssay(seu) <- "ADT"
  seu <- NormalizeData(seu, normalization.method = "CLR")
  seu <- ScaleData(seu)
  DefaultAssay(seu) <- "RNA"
  seu
}

## The continuous expression palette: ArchR::paletteContinuous(n=8)[c(1:3,6:8)].
bd_rna_cols <- function() {
  ArchR::paletteContinuous(n = 8)[c(1:3, 6:8)]
}

bd_clus_cols36 <- c(
  "#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b",
  "#e377c2", "#7f7f7f", "#bcbd22", "#17becf", "#aec7e8", "#ffbb78",
  "#98df8a", "#ff9896", "#c5b0d5", "#c49c94", "#f7b6d2", "#c7c7c7",
  "#dbdb8d", "#9edae5", "#ad494a", "#8c6d31", "#e7ba52", "#17becf",
  "#1f78b4", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#cab2d6",
  "#6a3d9a", "#ff8c00", "#b15928", "#41ab5d", "#f03b20", "#807dba")

bd_theme <- function() {
  library(ggplot2)
  theme_bw(base_size = 14) +
    theme(panel.background   = element_rect(fill = "transparent", colour = NA),
          panel.grid.minor   = element_blank(),
          panel.grid.major   = element_blank(),
          legend.background  = element_rect(fill = "transparent"),
          legend.box.background = element_rect(fill = "transparent"),
          legend.key         = element_rect(fill = "transparent", colour = NA),
          plot.background    = element_rect(fill = "transparent", colour = NA))
}

## Write the R and package versions beside the results.
bd_session <- function(path) {
  writeLines(c(capture.output(sessionInfo()), "",
               paste("run at", format(Sys.time(), "%Y-%m-%d %H:%M:%S"))), path)
}
```


## 1. load Cart1 + Cart2

```{r sec1-load-cart1-cart2}
dirs <- c(cart1 = file.path(DATA, "Cart1"),
          cart2 = file.path(DATA, "Cart2"))
loaded <- bd_load(dirs, which = "RSEC", n_pabo = 7)
seu <- bd_seurat(loaded)
cat("   cells before filtering:", ncol(seu), "\n")
```

## 2. QC filter, then re-process
```{r sec2-qc-filter-then-re-process}

seu <- bd_process(seu, dims = 1:40, resolution = 0.5, seed = SEED)
cat("   first pass:", ncol(seu), "cells,",
    nlevels(seu$seurat_clusters), "clusters at resolution 0.5\n")
seu <- subset(seu, subset = nCount_RNA > 1000 & nCount_RNA < 50000 & percent.mt < 40)
cat("   cells after filtering :", ncol(seu), "   (expected 10,363 from 30,391)\n")
seu <- bd_process(seu, dims = 1:40, resolution = 0.5, seed = SEED)
seu <- bd_process_adt(seu)
```

## demultiplex the four CD326 oligos
```{r demultiplex-the-four-cd326-oligos}
ab_cols <- c(Ab1 = "CD326",
             Ab2 = "CD326:G8.8-AMM2281",
             Ab3 = "CD326:G8.8-AMM2294",
             Ab4 = "CD326:G8.8-AMM2295")
group_of <- c(Ab1 = "Female_IFNGRpos", Ab2 = "Female_IFNGRneg",
              Ab3 = "Male_IFNGRpos",   Ab4 = "Male_IFNGRneg")

adt <- GetAssayData(seu, assay = "ADT", layer = "data")
stopifnot(all(ab_cols %in% rownames(adt)))
pos <- sapply(ab_cols, function(a) adt[a, ] > 1)     # Rmd threshold: CLR value > 1
colnames(pos) <- names(ab_cols)

## density of each tag - the plot behind the threshold choice
for (nm in names(ab_cols)) {
  d <- data.frame(value = adt[ab_cols[[nm]], ])
  pdfout(paste0("[QC] ADT density ", nm, " ", ab_cols[[nm]]), 5, 4)
  print(ggplot(d, aes(x = value)) + geom_density() +
        geom_vline(xintercept = 1, linetype = 2) +
        labs(title = paste0(nm, " = ", ab_cols[[nm]], " -> ", group_of[[nm]]),
             x = "CLR-normalised ADT"))
  dev.off()
}

n_tag <- rowSums(pos)
seu$multiplet    <- n_tag > 1
seu$Ab_not_found <- n_tag < 1
cat(sprintf("   multiplets: %d   no-tag: %d   singlets: %d\n",
            sum(seu$multiplet), sum(seu$Ab_not_found), sum(n_tag == 1)))
pdfout("[QC] multiplet"); print(DimPlot(seu, group.by = "multiplet")); dev.off()
pdfout("[QC] Ab_not_found"); print(DimPlot(seu, group.by = "Ab_not_found")); dev.off()

seu <- seu[, !seu$multiplet & !seu$Ab_not_found]
cat("   cells after de-multiplexing:", ncol(seu), "   (Rmd recorded 9,135 from 10,363)\n")
seu$group <- unname(group_of[colnames(pos)[apply(pos[Cells(seu), , drop = FALSE], 1, which.max)]])
print(table(seu$group, seu$cart))
pdfout("[QC] groups"); print(DimPlot(seu, group.by = "group")); dev.off()

saveRDS(seu, file.path(CDS, "all_cells_processed.RDS"))
```

## Figure S10

```{r figure-s10}
DefaultAssay(seu) <- "ADT"
pdfout("[FigS10A] QC Clusters")
print(DimPlot(seu, label = TRUE, label.size = 5, repel = TRUE, label.box = TRUE) +
      scale_fill_manual(values = rep("white", nlevels(seu$seurat_clusters)))); dev.off()

DefaultAssay(seu) <- "RNA"
for (g in c("Lgr5", "Muc2", "Trpm5")) {
  pdfout(paste0("[FigS10C] QC ", g)); print(FeaturePlot_scCustom(seu, features = g)); dev.off()
}

mean_lgr5 <- tapply(FetchData(seu, "Lgr5")[, 1], seu$seurat_clusters, mean)
mean_muc2 <- tapply(FetchData(seu, "Muc2")[, 1], seu$seurat_clusters, mean)
ord <- names(sort(mean_lgr5))
gapcut <- function(v) { sv <- sort(v); g <- diff(sv); if (!length(g)) return(Inf)
                        sv[which.max(g)] + max(g) / 2 }
goblet  <- names(mean_muc2)[mean_muc2 > gapcut(mean_muc2)]
lowest  <- ord[1]
cat("   Lgr5 rank (low -> high):", paste(ord, collapse = " "), "\n")
cat("   Muc2-high (goblet)     :", paste(sort(goblet), collapse = ","),
    "   (published: 5,6,8)\n")
cat("   lowest-Lgr5 cluster    :", lowest, "   (published: 9)\n")

as_published <- setequal(goblet, c("5","6","8")) && identical(lowest, "9")
if (as_published) {
  cat("   cluster numbering matches the published object; using 8 vs 5/6/9\n")
  ref <- "8"; others <- c("5","6","9")
} else {
  warning("cluster numbering does NOT match the published object - ",
          "S10B comparisons are derived, not the published ones")
  cat("   *** numbering differs; falling back to measured rule ***\n")
  ref <- ord[4]; others <- ord[1:3]
}
cat("   S10B comparisons: ", ref, " vs ", paste(others, collapse = "/"), "\n", sep = "")
my_comparisons <- lapply(others, function(x) c(x, ref))

pdfout("[FigS10B] QC Lgr5 Exp Level")
print(VlnPlot(seu, features = "Lgr5", pt.size = 0) +
      stat_summary(geom = "point", size = 10, colour = "black", shape = 95) +
      stat_compare_means(comparisons = my_comparisons, method = "wilcox.test") +
      ylim(0, 7)); dev.off()

mm <- FetchData(seu, c("Lgr5","Muc2","Trpm5"))
mm$cluster <- seu$seurat_clusters
mm <- mm %>% tidyr::pivot_longer(-cluster, names_to = "gene", values_to = "expr") %>%
  group_by(cluster, gene) %>% summarise(mean = mean(expr), .groups = "drop")
write.csv(mm, file.path(OUT, "mean_Lgr5_Muc2_Trpm5_by_cluster.csv"), row.names = FALSE)

pdfout("[FigS10D] DotPlot Lgr5 Muc2 Trpm5", 6, 6)
print(DotPlot(seu, features = c("Lgr5","Muc2","Trpm5")) + RotatedAxis()); dev.off()

write.csv(data.frame(cluster = names(mean_lgr5), mean_Lgr5 = as.numeric(mean_lgr5)),
          file.path(OUT, "mean_Lgr5_by_cluster.csv"), row.names = FALSE)
write.csv(as.data.frame(table(cluster = seu$seurat_clusters, group = seu$group)),
          file.path(OUT, "cluster_by_group.csv"), row.names = FALSE)
bd_session(file.path(OUT, "sessionInfo.txt"))
cat("== done ->", OUT, "\n")
```
