---
title: "Exp649 - Figure 7 and Figure S11"
author: "Koyama et al., Science Immunology"
output: html_document
---

# Exp649 — Figure 7 and Figure S11
# The Lgr5+ intestinal stem cell subset. Input is `all_cells_processed.RDS`, written by `01_figS10_qc.Rmd` — knit that file first.


```{r setup}
suppressPackageStartupMessages({
  library(Seurat); library(ggplot2); library(dplyr); library(tibble)
  library(scCustomize); library(ggpubr); library(Matrix)
  library(presto); library(fgsea); library(msigdbr)
})
```

## Setting up

**Edit one line — `DATA`.** Download this experiment's processed matrices
from NCBI GEO [GSE348009](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE348009)
to wherever you like and point `DATA` at that folder. It must contain
nothing - this file reads the `.RDS` written by `01_figS10_qc.Rmd`, so knit that first.

Output is written to
`Exp649_fig7_figS11/` in the working directory.

```{r paths}
## ==========  EDIT THIS ONE LINE  ==========
DATA <- "~/GSE348009/Exp649"
## ==========================================

IN  <- file.path("Exp649_figS10", "cds", "all_cells_processed.RDS")
OUT <- "Exp649_fig7_figS11"
FIG <- file.path(OUT, "figs"); CDS <- file.path(OUT, "cds")
for (d in c(FIG, CDS)) dir.create(d, recursive = TRUE, showWarnings = FALSE)

theme_set(bd_theme()); rna_cols <- bd_rna_cols(); clus_cols36 <- bd_clus_cols36
SEED <- 1234
pdfout <- function(n, w = 6, h = 6) {
  pdf(file.path(FIG, paste0(n, ".pdf")), width = w, height = h)
}

cat("== Exp649 Fig 7 / S11 ==\n")
seu <- readRDS(IN)
cat("   cells in:", ncol(seu), " clusters:", nlevels(seu$seurat_clusters), "\n")
```

## Helper functions
```{r functions}
suppressPackageStartupMessages({
  library(Matrix)
  library(data.table)
  library(Seurat)
})

## LogNormalize -> 3000 variable features -> scale -> PCA(50) -> neighbours,
## clusters, UMAP.
bd_process <- function(seu, dims, resolution, seed = 1234) {
  set.seed(seed)
  DefaultAssay(seu) <- "RNA"
  seu <- NormalizeData(seu, normalization.method = "LogNormalize", scale.factor = 10000)
  seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 3000)
  seu <- ScaleData(seu)
  seu <- RunPCA(seu, features = VariableFeatures(object = seu), npcs = 50)
  seu <- FindNeighbors(seu, dims = dims)
  seu <- FindClusters(seu, resolution = resolution)
  seu <- RunUMAP(seu, dims = dims)
  seu
}

## CLR normalisation of the protein assay.
bd_process_adt <- function(seu) {
  DefaultAssay(seu) <- "ADT"
  seu <- NormalizeData(seu, normalization.method = "CLR")
  seu <- ScaleData(seu)
  DefaultAssay(seu) <- "RNA"
  seu
}

## The continuous expression palette: ArchR::paletteContinuous(n=8)[c(1:3,6:8)].
bd_rna_cols <- function() {
  if (!requireNamespace("ArchR", quietly = TRUE))
    stop("ArchR is not installed; it defines rna_cols (paletteContinuous). ",
         "Run common/scripts/install_r_deps.sh first - do not substitute another palette.")
  ArchR::paletteContinuous(n = 8)[c(1:3, 6:8)]
}

bd_clus_cols36 <- c(
  "#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b",
  "#e377c2", "#7f7f7f", "#bcbd22", "#17becf", "#aec7e8", "#ffbb78",
  "#98df8a", "#ff9896", "#c5b0d5", "#c49c94", "#f7b6d2", "#c7c7c7",
  "#dbdb8d", "#9edae5", "#ad494a", "#8c6d31", "#e7ba52", "#17becf",
  "#1f78b4", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#cab2d6",
  "#6a3d9a", "#ff8c00", "#b15928", "#41ab5d", "#f03b20", "#807dba")

bd_theme <- function() {
  library(ggplot2)
  theme_bw(base_size = 14) +
    theme(panel.background   = element_rect(fill = "transparent", colour = NA),
          panel.grid.minor   = element_blank(),
          panel.grid.major   = element_blank(),
          legend.background  = element_rect(fill = "transparent"),
          legend.box.background = element_rect(fill = "transparent"),
          legend.key         = element_rect(fill = "transparent", colour = NA),
          plot.background    = element_rect(fill = "transparent", colour = NA))
}

## Write the R and package versions beside the results.
bd_session <- function(path) {
  writeLines(c(capture.output(sessionInfo()), "",
               paste("run at", format(Sys.time(), "%Y-%m-%d %H:%M:%S"))), path)
}
```





## ISC selection
```{r isc-selection}
DefaultAssay(seu) <- "RNA"
cl <- seu$seurat_clusters

## Lgr5+ ISC clusters: exclude Muc2+ goblet (5, 6, 8), Trpm5+ tuft (10)
## and the lowest-Lgr5 cluster (9).
keep <- c("0","1","2","3","4","7","11")

seu <- seu[, cl %in% keep]
cat("   ISC cells:", ncol(seu), "\n")

## re-embed the ISC subset (Rmd: dims 1:25, resolution 0.5)
seu <- bd_process(seu, dims = 1:25, resolution = 0.5, seed = SEED)
seu <- bd_process_adt(seu)
saveRDS(seu, file.path(CDS, "isc_processed.RDS"))
cat("   ISC clusters:", nlevels(seu$seurat_clusters), "\n")
```


## Figure 7B/D
```{r figure-7b-d}
pdfout("[Fig7B] Groups", 7, 6)
print(DimPlot_scCustom(seu, group.by = "group", colors_use = clus_cols36, pt.size = 1.5)); dev.off()

pdfout("[Fig7D] Clusters")
print(DimPlot_scCustom(seu, colors_use = clus_cols36, pt.size = 1.5, label = TRUE,
                       label.size = 8, repel = FALSE, label.box = TRUE) +
      scale_fill_manual(values = rep("white", nlevels(seu$seurat_clusters)))); dev.off()

for (g in c("Female_IFNGRneg","Female_IFNGRpos","Male_IFNGRneg","Male_IFNGRpos")) {
  seu[[paste0("group_", g)]] <- seu$group == g
  sub <- seu[, seu$group == g]
  pdfout(paste0("[Fig7B] group_", g))
  print(DimPlot_scCustom(sub, group.by = paste0("group_", g), colors_use = "Gray", pt.size = 1))
  dev.off()
}
```



## Figure 7E / S11A - MHC-II genes and protein
```{r figure-7e-s11a-mhc-ii-genes-and-protein}
options(warn = -1)
DefaultAssay(seu) <- "RNA"
for (g in c("Ciita","H2-Aa","H2-Ab1")) {
  pdfout(paste0("[Fig7E] ", g))
  print(FeaturePlot_scCustom(seu, features = g, label = FALSE, pt.size = 2,
                             max.cutoff = "q99") + scale_color_gradientn(colors = rna_cols) & NoAxes())
  dev.off()
}
DefaultAssay(seu) <- "ADT"
pdfout("[Fig7E] I-A-I-E")
print(FeaturePlot_scCustom(seu, features = "I-A-I-E", label = FALSE, pt.size = 2,
                           max.cutoff = "q99") + scale_color_gradientn(colors = rna_cols) & NoAxes())
dev.off()

## by cluster - the legend's clusters 4/5/6 comparisons
cmp_cl <- list(c("4","5"), c("5","6"), c("4","6"))
for (f in c("Ciita","H2-Aa","H2-Ab1")) {
  DefaultAssay(seu) <- "RNA"
  pdfout(paste0("[FigS11B] ", f, " Exp Level"))
  p <- VlnPlot(seu, features = f, pt.size = 0, cols = clus_cols36) +
       stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) + ylim(0, 7)
  p <- p + stat_compare_means(comparisons = cmp_cl, method = "wilcox.test")
  print(p); dev.off()
}
DefaultAssay(seu) <- "ADT"
pdfout("[FigS11B] I-A-I-E Exp Level")
p <- VlnPlot(seu, features = "I-A-I-E", pt.size = 0, cols = clus_cols36) +
     stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) + ylim(0, 7)
p <- p + stat_compare_means(comparisons = cmp_cl, method = "wilcox.test")
print(p); dev.off()

## by group - S11A
cmp_gp <- list(c("Female_IFNGRneg","Female_IFNGRpos"), c("Male_IFNGRneg","Male_IFNGRpos"),
               c("Female_IFNGRneg","Male_IFNGRneg"),   c("Female_IFNGRpos","Male_IFNGRpos"),
               c("Female_IFNGRneg","Male_IFNGRpos"),   c("Female_IFNGRpos","Male_IFNGRneg"))
for (f in c("Ciita","H2-Aa","H2-Ab1")) {
  DefaultAssay(seu) <- "RNA"
  pdfout(paste0("[FigS11A] ", f, " (by group)"))
  print(VlnPlot(seu, features = f, group.by = "group", pt.size = 0) +
        stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
        stat_compare_means(comparisons = cmp_gp, method = "wilcox.test") + ylim(0, 7)); dev.off()
}
DefaultAssay(seu) <- "ADT"
pdfout("[FigS11A] I-A-I-E (by group)")
print(VlnPlot(seu, features = "I-A-I-E", group.by = "group", pt.size = 0) +
      stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
      stat_compare_means(comparisons = cmp_gp, method = "wilcox.test") + ylim(0, 7)); dev.off()

## GOCC MHC class II module score by group
DefaultAssay(seu) <- "RNA"
m_cc <- msigdbr(species = "Mus musculus", category = "C5", subcategory = "GO:CC") %>%
  dplyr::select(gs_name, gene_symbol)
bac <- m_cc[m_cc$gs_name == "GOCC_MHC_CLASS_II_PROTEIN_COMPLEX", ]$gene_symbol
set.seed(SEED); seu <- AddModuleScore(seu, features = list(bac), name = "GOCC_MHC_CLASSII")
pdfout("[FigS11A] GOCC_MHC_CLASS_II_PROTEIN_COMPLEX Plot")
print(FeaturePlot_scCustom(seu, features = "GOCC_MHC_CLASSII1", label = FALSE, pt.size = 2,
                           max.cutoff = "q99") + scale_color_gradientn(colors = rna_cols) & NoAxes()); dev.off()
pdfout("[FigS11A] GOCC_MHC_CLASS_II_PROTEIN_COMPLEX Violin")
print(VlnPlot(seu, features = "GOCC_MHC_CLASSII1", group.by = "group", pt.size = 0) +
      stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
      stat_compare_means(comparisons = cmp_gp, method = "wilcox.test") + ylim(0, 4)); dev.off()
```


## Hallmark module scores (Fig 7B inset, S11)
```{r hallmark-module-scores}
m_h <- msigdbr(species = "Mus musculus", category = "H") %>% dplyr::select(gs_name, gene_symbol)
hall <- c(H_IFN = "HALLMARK_INTERFERON_GAMMA_RESPONSE")
for (nm in names(hall)) {
  genes <- m_h[m_h$gs_name == hall[[nm]], ]$gene_symbol
  set.seed(SEED); seu <- AddModuleScore(seu, features = list(genes), name = nm)
  pdfout(paste0("[Fig7B-S11] ", nm))
  print(FeaturePlot_scCustom(seu, features = paste0(nm, "1"), label = FALSE, pt.size = 2,
                             max.cutoff = "q99") + scale_color_gradientn(colors = rna_cols) & NoAxes())
  dev.off()
}
```

## fgsea bar graphs (Fig 7C, 7F)
```{r fgsea-bar-graphs}
plot_fsgea_bargraph <- function(genedb, genelist_wilcox, cluster, dbname,
                                NES_cutoff, q_value, height = 6, width = 6) {
  set.seed(SEED)
  fgsea_sets <- genedb %>% split(x = .$gene_symbol, f = .$gs_name)
  genes <- genelist_wilcox %>%
    mutate(auc = auc - 0.5) %>%
    group_by(feature) %>%
    dplyr::filter(group == cluster) %>%
    arrange(desc(auc)) %>%
    dplyr::select(feature, auc)
  ranks <- deframe(genes)
  fgseaRes <- fgsea(fgsea_sets, stats = ranks, nperm = 100000, nproc = 8)
  fgseaResTidy <- fgseaRes %>% as_tibble() %>% arrange(desc(NES))
  write.csv(do.call(cbind, fgseaResTidy[, 1:5]),
            file.path(OUT, paste0("GSEA - ", dbname, " - Group ", cluster, ".csv")))
  fgseaResTidy$pathway <- gsub("HALLMARK_", "", fgseaResTidy$pathway)
  pdf(file.path(FIG, paste0(dbname, "_Gp", cluster,
                            " q", q_value, " NES", NES_cutoff, ".pdf")), width, height)
  print(ggplot(fgseaResTidy[fgseaResTidy$padj < q_value & abs(fgseaResTidy$NES) > NES_cutoff, ],
               aes(reorder(pathway, NES), NES)) +
        geom_col(aes(fill = ES > 0)) + coord_flip() +
        labs(x = "Pathway", y = "Normalized Enrichment Score",
             title = paste0("q<", q_value, ", |NES|>", NES_cutoff)) + theme_classic())
  dev.off()
  invisible(ranks)
}

DefaultAssay(seu) <- "RNA"
## Fig 7C: male IFNGRpos vs male IFNGRneg
g_mm <- wilcoxauc(seu, "group", groups_use = c("Male_IFNGRpos", "Male_IFNGRneg"))
plot_fsgea_bargraph(m_h, g_mm, "Male_IFNGRpos", "[Fig7C] Hallmark MvM IFNGRneg v pos", 0, 0.2)

## Fig 7F and S11C: clusters 5 vs 6, and 4 vs 5, 4 vs 6
for (pair in list(c("4","5"), c("4","6"), c("5","6"))) {
  g <- wilcoxauc(seu, "seurat_clusters", groups_use = pair)
  for (side in pair)
    plot_fsgea_bargraph(m_h, g, side, paste0("[Fig7F-S11] Hallmark Set ", pair[1], "v", pair[2]), 0, 0.2)
}

write.csv(as.data.frame(table(cluster = seu$seurat_clusters, group = seu$group)),
          file.path(OUT, "cluster_by_group.csv"), row.names = FALSE)
saveRDS(seu, file.path(CDS, "isc_scored.RDS"))
bd_session(file.path(OUT, "sessionInfo.txt"))
cat("== done ->", OUT, "\n")
```
