---
title: "Exp649 - Figure 7 and Figure S11"
author: "Koyama et al., Science Immunology (ady3001)"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: show
---

<!-- GENERATED from 02_fig7_figS11.R by Zenodo/manual_build/r_to_rmd.py.
     Every line of R below is byte-identical to that script, which is the
     version verified to reproduce the published objects cell-for-cell.
     Re-generate rather than hand-editing, or re-verify if you do edit. -->

Exp649 - Figure 7 and Figure S11  (Lgr5+ ISC subset)

Source: Exp649/rmd/Exp649_Final.Rmd, chunks "- isolate ISCs" (second half),
  "- fsgea function", "Figure 6 Paper", "Supp Figure 1 Paper",
  "GSEA Figures / Supp Figure 2 Paper", "Supp Figure 3 Paper".
  (The Rmd's "Figure 6" is Figure 7 in the accepted manuscript, and its
   "Supp Figure 1/2/3" are S11; the numbering shifted during revision.)

Input is the object written by 01_figS10_qc.R.

ISC selection, exactly as the Fig S10 legend states it: drop the Muc2+ goblet
clusters (5, 6, 8), the Trpm5+ tuft cluster (10) and the lowest-Lgr5 cluster
(9), keeping 0, 1, 2, 3, 4, 7, 11 - which is what the Rmd hardcodes.

The rule is applied by MEASUREMENT and then checked against those numbers. In
practice the rebuild reproduces the published clustering cluster-for-cluster
(all twelve sizes identical), so the two agree; but if a future run renumbers,
this prints the disagreement instead of silently subsetting the wrong cells.

Usage: Rscript 02_fig7_figS11.R [--source deposit|original|rebuilt] [--which RSEC|DBEC]

> **Knitting this file.** The code locates the package root by walking up
> for a `.ady3001_root` marker, so knit with `knit_root_dir` set to the
> package root, or run from a directory beneath it. Point `$ADY3001_DATA`
> at an unpacked GEO download; `tools/unpack_geo_download.sh` builds that
> tree. Results are written to `$ADY3001_OUT`, default `./ady3001_output`.

<!-- original shebang: `#!/usr/bin/env Rscript` -->

```{r setup}
suppressPackageStartupMessages({
  library(Seurat); library(ggplot2); library(dplyr); library(tibble)
  library(scCustomize); library(ggpubr); library(Matrix)
  library(presto); library(fgsea); library(msigdbr)
})
```

## path resolution (Zenodo package)

These lines replace the hardcoded lab paths of the working copy. Every other
change to this file is also path-only; no analysis line differs from the
script that produced the published figures. The complete diff against the
working copy is shipped at 05_verification/path_rewrite.diff - read it rather
than taking this comment's word for it.

```{r path-resolution}
local({
  a <- commandArgs(FALSE)
  f <- sub("^--file=", "", a[grep("^--file=", a)])
  d <- if (length(f)) dirname(normalizePath(f[1])) else normalizePath(getwd())
  repeat {
    if (file.exists(file.path(d, ".ady3001_root"))) break
    p <- dirname(d)
    if (p == d) stop("cannot locate the package root (.ady3001_root marker)")
    d <- p
  }
  assign("PKG", d, envir = .GlobalEnv)
})
source(file.path(PKG, "02_figure_scripts", "common", "config.R"))

args <- commandArgs(trailingOnly = TRUE)
getopt <- function(f, d) { i <- match(f, args); if (is.na(i)) d else args[i + 1L] }
SOURCE <- getopt("--source", "deposit"); WHICH <- getopt("--which", "RSEC")

IN  <- file.path(OUT_ROOT, "Exp649",
                 paste0("outS10_", SOURCE, "_", WHICH), "cds", "all_cells_processed.RDS")
OUT <- file.path(OUT_ROOT, "Exp649", paste0("outFig7_", SOURCE, "_", WHICH))
FIG <- file.path(OUT, "figs"); CDS <- file.path(OUT, "cds")
## Create and VERIFY. dir.create(showWarnings = FALSE) hides a failure, and a
## failure here is silent until the first pdf() call dies hundreds of lines later.
for (d in c(FIG, CDS)) {
  dir.create(d, recursive = TRUE, showWarnings = FALSE)
  if (!dir.exists(d)) stop("could not create output directory: ", d)
}
if (!file.exists(IN)) stop("run 01_figS10_qc.R first; missing ", IN)

theme_set(bd_theme()); rna_cols <- bd_rna_cols(); clus_cols36 <- bd_clus_cols36
SEED <- 1234
## A "/" in the name would make pdf() try to write into a non-existent
## subdirectory, which fails only when that panel is reached. Reject it here.
pdfout <- function(n, w = 6, h = 6) {
  if (grepl("/", n, fixed = TRUE)) stop("panel name contains '/': ", n)
  pdf(file.path(FIG, paste0(n, ".pdf")), width = w, height = h)
}

cat("== Exp649 Fig 7 / S11 ==\n   source:", SOURCE, "\n")
seu <- readRDS(IN)
cat("   cells in:", ncol(seu), " clusters:", nlevels(seu$seurat_clusters), "\n")
```

## ISC selection

```{r isc-selection}
DefaultAssay(seu) <- "RNA"
x <- FetchData(seu, c("Lgr5", "Muc2", "Trpm5"))
cl <- seu$seurat_clusters
prof <- data.frame(cluster = levels(cl),
                   Lgr5  = as.numeric(tapply(x$Lgr5,  cl, mean)),
                   Muc2  = as.numeric(tapply(x$Muc2,  cl, mean)),
                   Trpm5 = as.numeric(tapply(x$Trpm5, cl, mean)))
## Goblet / Tuft clusters stand out by an order of magnitude; use a gap rule on
## the sorted means rather than an absolute number, so it does not depend on
## normalisation scale. The chosen sets are printed for checking.
gapcut <- function(v) { s <- sort(v); g <- diff(s); if (!length(g)) return(Inf)
                        s[which.max(g)] + max(g) / 2 }
goblet <- prof$cluster[prof$Muc2  > gapcut(prof$Muc2)]
tuft   <- prof$cluster[prof$Trpm5 > gapcut(prof$Trpm5)]
lowest <- prof$cluster[which.min(prof$Lgr5)]
drop   <- unique(c(goblet, tuft, lowest))
keep   <- setdiff(prof$cluster, drop)
PUBLISHED_KEEP <- c("0","1","2","3","4","7","11")
cat("   Muc2-high (goblet):", paste(sort(goblet), collapse = ","), "   (published: 5,6,8)\n")
cat("   Trpm5-high (tuft) :", paste(sort(tuft),   collapse = ","), "   (published: 10)\n")
cat("   lowest Lgr5       :", lowest, "   (published: 9)\n")
cat("   KEPT              :", paste(sort(keep), collapse = ","),
    "   (Rmd kept 0,1,2,3,4,7,11)\n")
if (setequal(keep, PUBLISHED_KEEP)) {
  cat("   selection matches the published ISC set\n")
} else {
  warning("ISC selection differs from the published set (", paste(keep, collapse=","),
          " vs ", paste(PUBLISHED_KEEP, collapse=","), ") - figures below are NOT ",
          "a reproduction of Fig 7 / S11")
  cat("   *** SELECTION DIFFERS FROM THE PUBLISHED SET ***\n")
}
write.csv(prof, file.path(OUT, "cluster_profile_Lgr5_Muc2_Trpm5.csv"), row.names = FALSE)
writeLines(c(paste("goblet:", paste(goblet, collapse = ",")),
             paste("tuft:",   paste(tuft,   collapse = ",")),
             paste("lowest_Lgr5:", lowest),
             paste("kept:",   paste(keep,   collapse = ",")),
             "Rmd kept: 0,1,2,3,4,7,11"),
           file.path(OUT, "isc_selection.txt"))

seu <- seu[, cl %in% keep]
cat("   ISC cells:", ncol(seu), "\n")

## re-embed the ISC subset (Rmd: dims 1:25, resolution 0.5)
seu <- bd_process(seu, dims = 1:25, resolution = 0.5, seed = SEED)
seu <- bd_process_adt(seu)
saveRDS(seu, file.path(CDS, "isc_processed.RDS"))
cat("   ISC clusters:", nlevels(seu$seurat_clusters), "  (published: 9)\n")
```

## Figure 7B/D

```{r figure-7b-d}
pdfout("[Fig7D] Clusters")
print(DimPlot_scCustom(seu, colors_use = clus_cols36, pt.size = 1.5, label = TRUE,
                       label.size = 8, repel = FALSE, label.box = TRUE) +
      scale_fill_manual(values = rep("white", nlevels(seu$seurat_clusters)))); dev.off()
pdfout("[Fig7B] Groups", 7, 6)
print(DimPlot_scCustom(seu, group.by = "group", colors_use = clus_cols36, pt.size = 1.5)); dev.off()

for (g in c("Female_IFNGRneg","Female_IFNGRpos","Male_IFNGRneg","Male_IFNGRpos")) {
  seu[[paste0("group_", g)]] <- seu$group == g
  sub <- seu[, seu$group == g]
  pdfout(paste0("[Fig7B] group_", g))
  print(DimPlot_scCustom(sub, group.by = paste0("group_", g), colors_use = "Gray", pt.size = 1))
  dev.off()
}
```

## Figure 7E / S11A - MHC-II genes and protein

```{r figure-7e-s11a-mhc-ii-genes-and-protein}
DefaultAssay(seu) <- "RNA"
for (g in c("Ciita","H2-Aa","H2-Ab1","Mki67")) {
  pdfout(paste0("[Fig7E] ", g))
  print(FeaturePlot_scCustom(seu, features = g, label = FALSE, pt.size = 2,
                             max.cutoff = "q99") + scale_color_gradientn(colors = rna_cols) & NoAxes())
  dev.off()
}
DefaultAssay(seu) <- "ADT"
pdfout("[Fig7E] I-A-I-E")
print(FeaturePlot_scCustom(seu, features = "I-A-I-E", label = FALSE, pt.size = 2,
                           max.cutoff = "q99") + scale_color_gradientn(colors = rna_cols) & NoAxes())
dev.off()

## by cluster - the legend's clusters 4/5/6 comparisons
lv <- levels(seu$seurat_clusters)
cl456 <- intersect(c("4","5","6"), lv)
cmp_cl <- if (length(cl456) == 3) list(c("4","5"), c("5","6"), c("4","6")) else
          { warning("clusters 4/5/6 not all present; per-cluster brackets skipped"); NULL }
for (f in c("Ciita","H2-Aa","H2-Ab1")) {
  DefaultAssay(seu) <- "RNA"
  pdfout(paste0("[FigS11B] ", f, " Exp Level"))
  p <- VlnPlot(seu, features = f, pt.size = 0, cols = clus_cols36) +
       stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) + ylim(0, 7)
  if (!is.null(cmp_cl)) p <- p + stat_compare_means(comparisons = cmp_cl, method = "wilcox.test")
  print(p); dev.off()
}
DefaultAssay(seu) <- "ADT"
pdfout("[FigS11B] I-A-I-E Exp Level")
p <- VlnPlot(seu, features = "I-A-I-E", pt.size = 0, cols = clus_cols36) +
     stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) + ylim(0, 7)
if (!is.null(cmp_cl)) p <- p + stat_compare_means(comparisons = cmp_cl, method = "wilcox.test")
print(p); dev.off()

## by group - S11A
cmp_gp <- list(c("Female_IFNGRneg","Female_IFNGRpos"), c("Male_IFNGRneg","Male_IFNGRpos"),
               c("Female_IFNGRneg","Male_IFNGRneg"),   c("Female_IFNGRpos","Male_IFNGRpos"),
               c("Female_IFNGRneg","Male_IFNGRpos"),   c("Female_IFNGRpos","Male_IFNGRneg"))
for (f in c("Ciita","H2-Aa","H2-Ab1")) {
  DefaultAssay(seu) <- "RNA"
  pdfout(paste0("[FigS11A] ", f, " (by group)"))
  print(VlnPlot(seu, features = f, group.by = "group", pt.size = 0) +
        stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
        stat_compare_means(comparisons = cmp_gp, method = "wilcox.test") + ylim(0, 7)); dev.off()
}
DefaultAssay(seu) <- "ADT"
pdfout("[FigS11A] I-A-I-E (by group)")
print(VlnPlot(seu, features = "I-A-I-E", group.by = "group", pt.size = 0) +
      stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
      stat_compare_means(comparisons = cmp_gp, method = "wilcox.test") + ylim(0, 7)); dev.off()

## GOCC MHC class II module score by group
DefaultAssay(seu) <- "RNA"
m_cc <- msigdbr(species = "Mus musculus", category = "C5", subcategory = "GO:CC") %>%
  dplyr::select(gs_name, gene_symbol)
bac <- m_cc[m_cc$gs_name == "GOCC_MHC_CLASS_II_PROTEIN_COMPLEX", ]$gene_symbol
set.seed(SEED); seu <- AddModuleScore(seu, features = list(bac), name = "GOCC_MHC_CLASSII")
pdfout("[FigS11A] GOCC_MHC_CLASS_II_PROTEIN_COMPLEX Plot")
print(FeaturePlot_scCustom(seu, features = "GOCC_MHC_CLASSII1", label = FALSE, pt.size = 2,
                           max.cutoff = "q99") + scale_color_gradientn(colors = rna_cols) & NoAxes()); dev.off()
pdfout("[FigS11A] GOCC_MHC_CLASS_II_PROTEIN_COMPLEX Violin")
print(VlnPlot(seu, features = "GOCC_MHC_CLASSII1", group.by = "group", pt.size = 0) +
      stat_summary(fun = median, geom = "point", size = 10, colour = "black", shape = 95) +
      stat_compare_means(comparisons = cmp_gp, method = "wilcox.test") + ylim(0, 4)); dev.off()
```

## Hallmark module scores (Fig 7B inset, S11)

```{r hallmark-module-scores}
m_h <- msigdbr(species = "Mus musculus", category = "H") %>% dplyr::select(gs_name, gene_symbol)
hall <- c(H_IFN = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
          H_IFNA = "HALLMARK_INTERFERON_ALPHA_RESPONSE",
          H_OXPHOS = "HALLMARK_OXIDATIVE_PHOSPHORYLATION",
          H_TNF = "HALLMARK_TNFA_SIGNALING_VIA_NFKB",
          H_TGFB = "HALLMARK_TGF_BETA_SIGNALING",
          H_E2F = "HALLMARK_E2F_TARGETS",
          H_G2M = "HALLMARK_G2M_CHECKPOINT",
          H_GLYCOLYSIS = "HALLMARK_GLYCOLYSIS",
          H_WNT = "HALLMARK_WNT_BETA_CATENIN_SIGNALING",
          H_HEDGEHOG = "HALLMARK_HEDGEHOG_SIGNALING")
for (nm in names(hall)) {
  genes <- m_h[m_h$gs_name == hall[[nm]], ]$gene_symbol
  if (!length(genes)) { warning("empty gene set ", hall[[nm]]); next }
  set.seed(SEED); seu <- AddModuleScore(seu, features = list(genes), name = nm)
  pdfout(paste0("[Fig7B-S11] ", nm))
  print(FeaturePlot_scCustom(seu, features = paste0(nm, "1"), label = FALSE, pt.size = 2,
                             max.cutoff = "q99") + scale_color_gradientn(colors = rna_cols) & NoAxes())
  dev.off()
}
```

## fgsea bar graphs (Fig 7C, 7F)

Verbatim transcription of plot_fsgeaTEST_bargraph() from the Rmd.

```{r fgsea-bar-graphs}
plot_fsgea_bargraph <- function(genedb, genelist_wilcox, cluster, dbname,
                                NES_cutoff, q_value, height = 6, width = 6) {
  set.seed(SEED)
  fgsea_sets <- genedb %>% split(x = .$gene_symbol, f = .$gs_name)
  genes <- genelist_wilcox %>%
    mutate(auc = auc - 0.5) %>%
    group_by(feature) %>%
    dplyr::filter(group == cluster) %>%
    arrange(desc(auc)) %>%
    dplyr::select(feature, auc)
  ranks <- deframe(genes)
  fgseaRes <- fgsea(fgsea_sets, stats = ranks, nperm = 1000000, nproc = 8)
  fgseaResTidy <- fgseaRes %>% as_tibble() %>% arrange(desc(NES))
  write.csv(do.call(cbind, fgseaResTidy[, 1:5]),
            file.path(OUT, paste0("GSEA - ", dbname, " - Group ", cluster, ".csv")))
  fgseaResTidy$pathway <- gsub("HALLMARK_", "", fgseaResTidy$pathway)
  pdf(file.path(FIG, paste0("[GSEA] ", dbname, "_Gp", cluster,
                            " q", q_value, " NES", NES_cutoff, ".pdf")), width, height)
  print(ggplot(fgseaResTidy[fgseaResTidy$padj < q_value & abs(fgseaResTidy$NES) > NES_cutoff, ],
               aes(reorder(pathway, NES), NES)) +
        geom_col(aes(fill = ES > 0)) + coord_flip() +
        labs(x = "Pathway", y = "Normalized Enrichment Score",
             title = paste0("q<", q_value, ", |NES|>", NES_cutoff)) + theme_classic())
  dev.off()
  invisible(ranks)
}

DefaultAssay(seu) <- "RNA"
## Fig 7C: male IFNGRpos vs male IFNGRneg
g_mm <- wilcoxauc(seu, "group", groups_use = c("Male_IFNGRpos", "Male_IFNGRneg"))
plot_fsgea_bargraph(m_h, g_mm, "Male_IFNGRpos", "Hallmark MvM IFNGRneg v pos", 0, 0.2)
plot_fsgea_bargraph(m_h, g_mm, "Male_IFNGRneg", "Hallmark MvM IFNGRpos v neg", 0, 0.2)

## Fig 7F and S11C: clusters 5 vs 6, and 4 vs 5, 4 vs 6
if (length(cl456) == 3) {
  for (pair in list(c("4","5"), c("4","6"), c("5","6"))) {
    g <- wilcoxauc(seu, "seurat_clusters", groups_use = pair)
    for (side in pair)
      plot_fsgea_bargraph(m_h, g, side, paste0("Hallmark Set ", pair[1], "v", pair[2]), 0, 0.2)
  }
} else warning("clusters 4/5/6 not all present; cluster GSEA skipped")

write.csv(as.data.frame(table(cluster = seu$seurat_clusters, group = seu$group)),
          file.path(OUT, "cluster_by_group.csv"), row.names = FALSE)
saveRDS(seu, file.path(CDS, "isc_scored.RDS"))
bd_session(file.path(OUT, "sessionInfo.txt"))
cat("== done ->", OUT, "\n")
```
